### 2019-06-15 misc/study-notes/ndmi011notes.tex
Added graph theory notes and TeX macros

% Notes for NDMI011 Combinatorics and Graph Theory
% Charles University in Prague, Summer 2019
% Instructor: Prof. Andreas Feldmann
% Notes by Marcel Goh

\input fontmac
\input mathmac

\leftrighttop{NDMI011 COMBINATORICS AND GRAPH THEORY}{MARCEL GOH}

\maketitle{NDMI011 Combinatorics and Graph Theory}{Notes by}{Marcel K. Goh {\rm (Prague, Czech Rep.)}}{10 June 2019}

\section 1. PRELIMINARIES

We say that a function $f(n)$ is $O\big(g(n)\big)$ (read ``big-oh of $g(n)$'') if there exist constants $n_0$ and $C$ such that for all $n\geq n_0$, $f(n)\leq C\cdot g(n)$. If the ratio ${f(n)\over g(n)}$ approaches 0 as n approaches infinity, then we say $f(n)$ is $o\big(g(n)\big)$ (read ``little-oh of $g(n)$''). If $f(n)$ is both $O\big(g(n)\big)$ and $o\big(g(n)\big)$, then $f(n)$ is $\Theta\big(g(n)\big)$ (read ``big-theta of $g\big(n)\big)$''). Finally, if the ratio ${f(n)\over g(n)}$ approaches 1 as $n$ approaches infinity, then we write $f(n) \sim g(n)$.

Let $G=(V,E)$ be a graph. The graph $G'=(V',E')$ is a {\it subgraph} of $G$ if $V'\subseteq V$ and $E'\subseteq E$. A {it cycle} is a graph on $v_1, \ldots, v_n$ with edges $\{v_1, v_2\}, \ldots, \{v_{n-1}, v_n\}$. A {\it tree} is a connected graph that contains no cycle (i.e\  does not have a cycle as a subgraph).

\proclaim Theorem. In any graph $G=(V,E)$,
$$\displaystyle\sum_{v\in V} \deg v = 2\cdot|E|$$

\proof Every edge $\{u, v\}\in E$ adds 1 to the degree of $u$ and adds 1 to the degree of $v$. Hence if you sum over all vertices, each edge gets counted twice.\slug

\proclaim Theorem.
The number of vertices with odd degree is even.

\proof Divide $V$ into $V_1$ and $V_2$ where $V_1$ is the set of vertices with odd degree and $V_2$ is the set of vertices with even degree. Then we have $\sum_{v\in V} \deg v = \sum_{v\in V_1} \deg v + \sum_{v\in V_2} \deg v$. Observe that the sum over all odd vertices $\sum_{v\in V_1}\deg v$ equals $2\cdot|E| - \sum_{v\in V_2}\deg v = 2(|E| - x)$ for some $x$, meaning it is even. Since the sum is even but each vertex has an odd degree, the number of vertices must be even.\slug

\proclaim Theorem.
A tree with $n$ vertices has $n-1$ edges.

\proof Let $G = (V,E)$ be a tree and let $n$ denote $|V|$, the number of vertices. Using induction on $n$, we show that it has $n-1$ vertices. The base case $n = 1$ is easy, since a tree with only one vertex has no edges. Now assume that a tree with $n$ vertices has $n-1$ edges and we consider the case where $G$ has $n+1$ vertices. Let us create $G'$ by removing a vertex from $G$. We cannot remove an internal vertex, since then $G'$ would not be connected. So we must remove a leaf, along with the edge that connected it to $G$. Now $G'$ has $n$ vertices, so by the induction hypothesis it has $n-1$ edges. This implies that $G$ had $n$ edges.
\slug

\section 2. GENERATING FUNCTIONS

\vskip -\medskipamount

\section 2.1. Power series and generating functions

Given an infinite sequence $a_0, a_1, \ldots = (a_i)_{i=0}^\infty$, the power series
$$\sum_{i=0}^\infty a_ix^i$$
is called the {\it generating function} of $(a_i)_{i=0}^\infty$ and is denoted $a(x)$. For example, the sequence $1, 1, 1, \ldots$ has generating function $\sum_{i=0}^\infty x^i$, which converges to $1/(1-x)$ for $x\in (-1, 1)$. Since we only really care about the coefficients of a generating function, we assume that $x$ is within the power series' interval of convergence.

\section 2.2. Binomial coefficients

For $r\in \R$ and $k\in \N$, the {\it binomial coefficient} is given by
$${r\choose k} = \left(\prod_{i=0}^{k-1} (r-i)\right) /k!$$
where ${r\choose 0} = 1$. If $r\in \N$, then we can think of ${r\choose k}$ as the number of ways to choose $k$ objects out of a pool of $r$ objects.

\thmparens{}{Generalised binomial theorem} For any $r\in \R$,
$$\sum_{i\geq 0} {r\choose i} x^i = (1 + x)^r.$$

\proof In this proof we will gloss over technicalities relating to convergence. Let $a(x)$ denote $(1+x)^r$. Then the derivative of $a(x)$ is $a'(x)=r(1+x)^{r-1}$, the second derivative $a''(x)$ is $r(r-1)(1+x)^{r-2}$, and so on. Generally,
$$a^{(i)}(x) = \left(\prod_{j=0}^{i-1} (r-j)\right)(1+x)^{r-i}.$$
If we evaluate the function at $x=0$, we get
$$a^{(i)}(0) = \prod_{j=0}^{i-1} (r-j)$$
so we can use Taylor's theorem to get
$$\eqalign{
    a(x) &= \sum_{i\geq 0} {a^{(i)}(0)\over i!} x^i\cr
    &= \sum_{i\geq 0}{r\choose i} x^i,
}$$
which is what we wanted.\slug

\section 2.3. Existence of a closed form

It is clear from the theorem above that the generating function for the sequence ${r\choose 1}, {r\choose 2}, \ldots$ is $(1+x)^r$. We may wonder what kinds of sequences have a {\it closed form} generating function. The following theorem, presented without proof, answers this question.

\proclaim Theorem. Let $(a_i)_{i=0}^\infty$ be a sequence of reals such that there exists some $k>0$ for which $|a_i|\leq k^i$ for every $i\geq 1$. Then for all $x$ in the interval $(-1/k, 1/k)$, the power series
$$a(x) = \sum_{i=0}^\infty a_ix^i$$
converges, so the generating function $a(x)$ has a closed form. Moreover, for any arbitrarily small $\epsilon >0$, the values of $a(x)$ for $x$ in the interval $(-\epsilon, \epsilon)$ uniquely determine the sequence $(a_i)_{i=0}^\infty$ where, by Taylor's theorem, $a_i = a^{(i)}(0)/i!$\slug

\section 2.4. Application to counting

Generating functions give us a method for counting the number of elements with some parameter $i$. For example, we could count the number of trees with $i$ vertices or the number of steps performed by an algorithm with input size $i$. Often, $a_i$ is defined inductively or recursively.

For example, suppose that for some number $c$ our sequence $a_i$ is given by $a_0 = c$, and $a_i = a_{i-1} + c$ for $i\geq 1$. Then we get that
$$\eqalign{
    a(x) &= \sum_{i\geq 0} a_ix^i\cr
    &= cx^0 + \sum_{i\geq 1} (a_{i-1} + c) x^i\cr
    &=  cx^0 + \sum_{i\geq 1}cx^i + \sum_{i\geq 1} a_{i-1}\cr
    &= {c\over 1-x} + x\cdot a(x).
}$$
This implies that $a(x) = c/(1-x)$, and by Taylor's theorem, $a_i = a^{(0)}(0)/i!$. Calculating this directly can be very tedious, so very often, we will express $a(x)$ in terms of generating functions that we know. In the case of this example, we note that
$$\eqalign{
    {c\over (1-x)^2} &= c\cdot {1\over 1-x}\cdot {1\over 1-x} \cr
    &= c\left(\sum_{i\geq 0} x^i\right) \left(\sum_{i\geq 0} x^i\right)\cr
    &= \sum_{i\geq 0} c(i+1) x^i,
}$$
which gives us that $a_i = c(i+1)$ (which we could have easily gotten by observing the recurrence).

\section 2.5. Operations on generating functions

Some useful operations on generating functions include:
\medskip
{\bf Addition.} $a(x) + b(x) = \displaystyle\sum_{i\geq 0} (a_i + b_i) x^i$.
\smallskip
{\bf Multiplication.} $\beta a(x) = \displaystyle\sum_{i\geq 0} (\beta a_i) x^i$.
\smallskip
{\bf Scaling.} $a(\beta x) = \displaystyle\sum_{i\geq 0} (\beta^i a_i) x^i$.
\smallskip
{\bf Gaps.} $a(x^k) = \displaystyle\sum_{i\geq 0} a_i x^{ki}$.
\smallskip
{\bf Shift right.} $xa(x) = \displaystyle\sum_{i\geq 0} (a_{i-1}) x^{i}$.
\smallskip
{\bf Shift left.} $\left(a(x) - a_0\right)/x = \displaystyle\sum_{i\geq 0} (a_{i+1}) x^{i}$.
\smallskip
{\bf Derivative.} $a'(x) = \displaystyle\sum_{i\geq 0} \big((a+1) a_{i+1}\big) x^{i}$.
\smallskip
{\bf Integral.} $\displaystyle\int_0^x a(x)dx = \displaystyle\sum_{i\geq 1} (a_{i-1}/i) x^{i}$.
\smallskip
{\bf Convolution.} $a(x)b(x) = \displaystyle\sum_{i\geq 0} \left(\displaystyle\sum_{k=0}^i a_k b_{i-k}\right) x^{i}$.
\smallskip
{\bf Partial sums.} $a(x)/(1-x) = \displaystyle\sum_{i\geq 0} \left(\displaystyle\sum_{k\geq 0} a_k \right) x^{i}$.
\medskip

\section 2.6. Some applications

Imagine a situation at a grocery store. A customer has six \$1 coins, five \$2 coins, and four \$5 bills. Suppose she has to pay \$21. How many ways can she do this? To find the answer, we need only compute the coefficient of $x^{21}$ in the polynomial
$$ (1 + x + x^2 + \cdots + x^6)\cdot(1+x^2+x^4+\cdots+x^{10})\cdot(1+x^5+x^{10}+\cdots+x^{20}).$$
Now suppose that a box contains 30 red balls, 40 blue balls, and 50 white balls. In how many ways can one draw 70 balls? As you may have suspected, we need to get the coefficient of $x^{70}$ in
$$ (1+x+\cdots+x^{30})\cdot(1+x+\cdots+x^{40})\cdot(1+x+\cdots+x^{50}).$$
It may be instructive to work through a full example. We will use generating functions to count binary trees. A {\it binary tree} $T$ is either empty ($T=\emptyset$) or consists of a root adjacent to a left subtree $T_l$ and a right subtree $T_r$. Let $b_n$ denote the number of binary trees with $n$ vertices. We will try to find a closed formula for $b_n$. There is exactly one binary tree with 0 vertices (the empty one), and one binary tree with 1 vertex. From the definition, $n=0$ or $n = 1 + n_l + n_r$ where $n_l = k$ and $n_r = n-1-k$ for some $k\in \{0,\ldots, n-1\}$. Hence we obtain the recurrence $b_n = b_0\cdot b_{n-1} + b_1\cdot b_{n-2} + \cdots b_{n-1}\cdot b_0$ for $n\geq 1$, i.e.
$$ b_n = \sum_{k=0}^{n-1} b_k\cdot b_{n-1-k}. $$
Then the generating function $b(x)$ can be derived:
$$\eqalign{
    b(x) &= \sum_{n\geq 0} b_nx^n\cr
    &= 1x^0 + \sum_{n\geq 1} \left(\sum_{k\geq 0}^{n-1} b_k\cdot b_{n-1-k}\right) x^n\cr
    &= 1 + x\sum_{n\geq 0} \left(\sum_{k\geq 0}^{n-1} b_k\cdot b_{n-1-k}\right) x^n\cr
    &= 1 + xb(x)b(x).
}$$
This implies that $xb(x)^2 - b(x) + 1 = 0$, which means that
$$b(x) = {1\pm\sqrt{1-4x}\over 2x}$$
Which one is correct? Well, note that $b(0) = \sum_{n\geq 0} b_n0^n = b_0 = 1$, whereas
$$\lim_{x\rightarrow 0} {1 + \sqrt{1-4x}\over 2x} = +\infty.$$
We try the other root, starting with
$$\lim_{x\rightarrow 0} {1 - \sqrt{1-4x}\over 2x},$$
upon which we can apply l'Hospital's rule to get a limit of 1. So this is the generating function we want.

Applying Newton's formula and the scaling operation, we observe that
$$\sqrt{1-4x}=\big(1+(-4x)\big)^{1/2} = \sum_{n\geq 0} (-4)^n {{1/2}\choose n}x^n.$$
We can substitute this into our generating function to get that
$$b(x) = {1\over 2x}\left( 1 - \sum_{n\geq 0} (-4)^n {{1/2}\choose n}x^n\right).$$
Since $1 = \sum_{n\geq 0} (-4)^n {{1/2}\choose n}x^0$, we can further simplify, obtaining
$$\eqalign{
    b(x) &= {1\over x}\sum_{n\geq 1} -{1\over 2}(-4)^n{{1/2}\choose n}x^n\cr
    &= \sum_{n\geq 0} -{1\over 2} (-4)^{n+1}{{1/2}\choose n+1}x^n\cr
    &= \sum_{n\geq 0} {1\over n+1}{2n\choose n}x^n.
}$$
So the number of binary trees with $n$ vertices is ${2n\choose n}/(n+1)$. These are the {\it Catalan numbers}.

\section 3. GRAPHS

\vskip -\medskipamount

\section 3.1. Network flow

\vskip -\medskipamount

\section 3.1.1. Edge capacities

Suppose we are given a directed graph $G=(V,E)$ ($E\subseteq V\times V$) with designated source $s\in V$ and target $t\in V$. To each edge $e$ is assigned a capacity $c(e)$, which may be a non-negative real number or $+\infty$. Then an {\it s-t-flow} is a function $f:E\rightarrow \R$ such that for every edge $e\in E$, $0\leq f(e)\leq c(e)$. These are called the {\it capacity constraints}. Our flow must also obey the law of {\it flow conservation}, also known as {\it Kirchhoff's First Law} (which was originally observed in electrical circuits): For every vertex $v\in V\setminus\{s, t\}$, $\sum_{uv\in E} f(uv) = \sum_{vu\in E} f(vu)$. We want to maximise the flow that leaves $s$, subject to these constraints. The {\it value} of an s-t-flow is
$$|f| = \sum_{su\in E} f(su) - f(us),$$
and a {\it max-flow} is an s-t-flow with maximum value.

The intuitive way to check that a flow is maximal is to study ``bottlenecks''. Concretely, if we let $S\subseteq V$ be such that $s\in S$ and $t\notin S$, then the set $\delta^+(S) = \{uv\in E : u\in S, v\notin S\}$ is an {\it s-t-cut} or {\it s-t-edge-cut}. After removing $\delta^+(S)$ from $G$, there is no longer a path from $s$ to $t$, i.e.\  $\delta^+(S)$ {\it separates} $s$ from $t$. (Since we only removed edges {\it leaving} $S$, there may still be a path from $t$ to $s$.) The {\it capacity} of an s-t-cut is given by
$$ c\big(\delta^+(S)\big) = \sum_{uv\in \delta^+(S)} c(uv).$$
Then we can define a {\it min-cut} to be an s-t-cut with minimum capacity.

\proclaim Lemma. For any s-t-flow $f$ and s-t-cut $\delta^+(S)$,
$$|f| = \sum_{e\in \delta^+(S)} f(e) - \sum_{e\in \delta^-(S)} f(e),$$
where $\delta^-(S)$ is the set $\{uv : u\notin S, v\in S\}$.

\proof By flow conservation, we know that for every $v \neq s,t$,
$$\sum_{vu\in E} f(vu) - \sum_{uv\in E} f(uv) = 0.$$
Then the value of the s-t-flow is the sum of all equations for the cut $S$:
$$|f| = \sum_{v\in S} \left( \sum_{vu\in E} f(vu) - \sum_{uv\in E} f(uv) \right).$$
We need to consider three possible cases for an edge $e = uv$:
\medskip
\item {1.} ($u, v\in S$) If both $u$ and $v$ are in $S$, then we will add $f(uv)$ and then subtract $f(uv)$, so this case does not contribute to $|f|$.
\smallskip
\item {2.} ($u\in S$, $v\notin S$) In this case, $uv\in \delta^+(S)$, so we will have a contribution of $+f(uv)$ to $|f|$.
\smallskip
\item {3.} ($u\notin S$, $v\in S$) In this case, $uv\in \delta^-(S)$, so we will have a contribution of $-f(uv)$ to $|f|$.
\medskip
This implies that
$$ |f| = \sum_{e\in \delta^+(S)} f(e) - \sum_{e\in \delta^-(S)} f(e) ,$$
which is exactly what we wanted to prove.\slug

We can use this to derive an easy corollary:

\proclaim Corollary. If $f$ is a max-flow and $\delta^+(S)$ is a min-cut, then $|f| \leq c\big(\delta^+(S)\big)$.

\proof By the preceding lemma, $|f| \leq \sum_{e\in \delta^+(S)} f(e)$, which by capacity constraints is no greater than $\sum_{e\in \delta^+(S)} c(e)$. But by the definition of the capacity of a min-cut, this is just $c\big(\delta^+(S)\big)$. \slug

In fact, the value of a max-flow {\it equals} the capacity of a min-cut, but the other inequality requires a more involved proof.

\proclaim Theorem. If $f$ is a max-flow and $\delta^+(S)$ is a min-cut, then $|f| = c\big(\delta^+(S)\big)$.

\proof By the preceding corollary, it suffices to show that $|f| \geq c\big(\delta^+(S)\big)$. For any max-flow $f$, the idea is to construct some s-t-cut $\delta^+(S)$ with $|f| = \delta^+(S)$. This would imply that any {\it min-cut} would have capacity no greater than $|f|$.

Consider the following procedure. We start with $S\gets \{s\}$. While there exists an edge $uv\in \delta^+(S)$ such that $c(uv) > f(uv)$ {\it or} there exists an edge $vu\in \delta^-(S)$ such that $f(uv) > 0$, we add $v$ to $S$. The claim is that if $f$ is a max-flow, then this algorithm will produce an $S$ such that $t\notin S$, i.e.\  $S$ is an s-t-cut. If this claim is true, then we are done; since $\sum_{e\in \delta^-(S)} = 0$ after the algorithm terminates, we will have $|f| = c\big(\delta^+(S)\big)$.

Suppose, towards a contradiction, that $t\in S$. Then there exists a sequence of vertices $v_0,\ldots, v_k\in S$ such that $v_0 = s$, $v_k = t$, and for all $i\in \{0,\ldots, k-1\}$, either
\medskip
\item {a)} $v_iv_{i+1}\in E$ and $c(v_iv_{i+1}) > f(v_iv_{i+1})$, or
\smallskip
\item {b)} $v_{i+1}v_i\in E$ and $f(v_{i+1}v_i) > 0$.
\medskip
For each $i$, define a small real number $\epsilon_i$ given by
$$\epsilon_i = \cases{
    c(v_iv_{i+1}) - f(v_iv_{i+1}) & \hbox{if case (a) holds}\cr
    f(v_{i+1}) & \hbox{if case (b) holds}
}$$
Then we can let $\epsilon = \min\{e_i : 0\leq i\leq k-1\}$. Note that, by construction, $\epsilon > 0$.

Now we create a new function $f':E\rightarrow \R$ given by
$$ f'(e) = \cases{
    f(e) + \epsilon & \hbox{if}\ e\ \hbox{satisfies case (a)}\cr
    f(e) - \epsilon & \hbox{if}\ e\ \hbox{satisfies case (b)}\cr
    f(e) & \hbox{otherwise}
}$$
The claim is that $f'$ is an s-t-flow. The values of $f'(e)$ are non negative, because $f'(e)\geq f(e)\geq 0$ in case a), and in case b), $f'(e) = f(e) - \epsilon \geq 0$. The capacity constraints are satisfied, since in case a), $\epsilon\leq c(e) - f(e)$ implies that $f'(e) = f(e) + \epsilon \leq c(e)$; and in case b), $f'(e)\leq f(e)\leq c(e)$. And we can see that the law of flow conservation is obeyed as well, since for every $v\in V\setminus\{s, t\}$, either flow is unchanged at $v$ or exactly two edges are changed (one adds $\epsilon$ and the other one subtracts $\epsilon$ from $f$).

Computing the value of the flow $f'$, we find that
$$ |f'| = \sum_{sv\in E} f'(sv) - \sum_{vs\in E} f'(vs) = |f| + \epsilon.$$
This implies that $f$ is not a max-flow, a contradiction.\slug

So to determine if a flow is maximal, we can figure out what $S$ is and see if $t\in S$. If not, then $f$ is a max-flow and $\delta^+(S)$ is a min-cut. The proof of the theorem also suggests an algorithm to compute max-flow and min-cut:

\algbegin Algorithm M (Find max-flow/min-cut). Given a directed graph $G=(V,E)$, this algorithm finds a max-flow $f$ and a set $S\subset V$ such that $\delta^+(S)$ is a min-cut.

\algstep M1. [Initialise.] Set $i\gets 0$ and $f_0(e) \gets 0$ for every $e\in E$.
\algstep M2. [Find $S_i$] Construct the set $S_i$ for the flow $f_i$ using the procedure outlined in the preceding proof.
\algstep M3. [Is $t$ in $S_i$?] If $t\notin S_i$, the flow $f_i$ is maximal. Output $f_i$ and $S_i$.
\algstep M4. [Increase flow.] Find the ``path''$P$ from $s$ to $t$ in $S$, and compute $\epsilon$. Let $f_{i+1}$ be the result of augmenting $f_i$ by $\epsilon$ along $P$. Set $i\gets i+1$.
\algstep M5. [Repeat.] Go to step M2.\slug

It is clear that if the algorithm terminates, it will output the correct answer. So it is natural to wonder when the algorithm is sure to terminate. In the case of integer capacities, termination is easy to prove:

\proclaim Theorem. If for all $e\in E$, $c(e)\in \N_0$, then the capacity $n$ of the min-cut is integral and Algorithm M terminates in at most $n$ iterations.

\proof We prove, by induction on $i$, that at every step of the algorithm, $|f_i|\in \N_0$. The base case is simple because the algorithm starts with $|f_0|=0$. Now assume that $f_i\in \N_0$. In the step that increases $i$ to $i+1$, we choose $\epsilon$ to be the minimum value computed over a path $P$ of edges. So for some $e\in P$, either $\epsilon = c(e) - f_i(e)$ or $\epsilon = f_i(e)$. Since both $c(e), f_i(e)\in \N_0$, we have that $\epsilon\in \N_0$ and the value of the flow at each iteration is integral.

Then because $\epsilon > 0$, we have from integrality that $\epsilon \geq 1$ meaning that for every iteration of the algorithm, the value $|f_{i+1}| = |f_i| + \epsilon \geq |f_i| + 1$. Since the value of the max-flow is exactly $n$, the algorithm terminates in at most $n$ iterations.\slug

\section 3.1.2. Vertex capacites

Instead of constraining flow at the edges, we can instead assign capacities $d(v)$ to vertices $v\in V\setminus\{s,t\}$. Then for all $v\in V\setminus\{s,t\}$, an s-t-flow $f:E\rightarrow \R$ must satisfy $\sum_{uv\in E} f(uv)\leq d(v)$ and $\sum_{uv\in E} f(uv)= \sum_{vu\in E} f(vu)$. Note that by this definition, if $st\in E$, then the value of the max-flow will be infinite, so assume that $st\notin E$.

Now we define an {\it s-t-vertex-cut} to be a set $T\subseteq V\setminus\{s,t\}$ such that the graph $G\setminus T$ has no path from $s$ to $t$. For ease of notation, let $d(T) = \sum_{v\in T} d(v)$.

\proclaim Theorem. Let $f$ be a max-flow in a network with vertex capacities. If $T$ is a min-cut, then $|f| = d(T)$, and furthermore, if all capacities are integers, then there is an integer max-flow.

\proof We simply convert our vertex-constrained network to an edge-constrained one. Replace each constrained vertex $v_i$ with two new vertices $v_{i1}$ and $v_{i2}$. Connect these vertices with a single edge $e_i$ with $c(e_i) = d(v_i)$. Set the capacties of all other edges in the new graph to $+\infty$. Then the theorem follows from our previous theorems regarding edge-constrained networks.\slug

\section 3.2. Bipartite matchings

A graph $G=(V,E)$ is {\it bipartite} if $V$ can be split into disjoint subsets $A$ and $B$ such that every edge in $E$ has one endpoint in $A$ and the other in $B$. A {\it matching} is a set $M\subseteq E$ such that no two edges of $M$ share a vertex. A {\it maximum matching} is matching of maximum size.

Consider the so-called {\it marriage problem}. Given a set $A$ of boys and $B$ of girls, can we marry off all boys to girls that they know? We can model this problem with a bipartite graph $(A\cup B, E)$, where an edge exists between a boy and a girl if they know each other. First we define the set of {\it neigbours} of a vertex set $S\subseteq A$ to be $N(S) = \{v\in B : uv\in E\ \hbox{and}\ u\in S\}$. Then the following theorem tells us exactly when the marriage problem is solvable.

\thmparens{}{Hall} A bipartite graph $G=(A\cup B, E)$ (with $A$ and $B$ disjoint) has a matching $M\subseteq E$ of size $|M| = |A|$ if and only if $|N(S)|\geq |S|$ for all $S\subseteq A$.

\proof The ``only if'' direction is obvious. We prove the ``if'' direction by contrapositive, i.e.\  if there is no matching of size $|A|$ then there exists some $S\subseteq A$ such that $|N(S) < |S|$.

The idea is to use the vertex version of the max-flow min-cut theorem. We construct a directed graph $H=(V',E')$ from G. Introduce two new vertices $s$ and $t$ and set $V' = A\cup B \cup \{s,t\}$. We want to connect $s$ to every vertex in $A$ and connect every vertex in $B$ to $t$. Then for every edge already in the graph, ensure it is directed from $A$ to $B$. So
$$ E' = \{ su : u\in A\}\cup\{vt : v\in B\}\cup\{uv : uv\in E, u\in A, v\in B\}.$$
Now we set all the capacities on every vertex $v\in A\cup B$ to 1. We know from the max-flow min-cut theorem that there exists an integer max-flow, call it $f$, which is a {\it 0-1 flow}, meaning that $f(e)\in \{0,1\}$ for all $e\in E'$. Note that if $f\geq |A|$, then there exists a matching of size $|A|$, since the 0-1 flow uses at most one incoming and one outgoing edge of any vertex in $V\setminus \{s,t\}$.

But we assumed that no matching of size $|A|$ exists, so $|f| < |A|$. Let $T$ be a min-cut. From the max-flow min-cut theorem, we know that $d(T) = |f| < |A|$. Consider the sets $X = A\cap T$ and $Y = B\cap T$. Since $|X| + |Y| = |T| = \sum_{v\in T} 1 = d(T)$, we have that $|Y| < |A| - |X|$. There is no edge from $A\setminus X$ to $B\setminus Y$, as $T$ is an s-t-vertex-cut. So $N(A\setminus X)\subseteq Y$ in $G$ and thus $|N(A\setminus X)| \leq |Y| < |A| - |X| = |A\setminus X|$. The set $A\setminus X$ is exactly the $S$ we were looking for.\slug

In a graph $G = (V,E)$, a {\it vertex cover} is a set $C\subseteq V$ such that every edge is incident to at least one vertex in $C$. A {\it minimum vertex cover} is a vertex cover of minimum cardinality. Notice that every vertex cover of $G$ is an s-t-vertex cut in $H$, otherwise there would be a path from $s$ to $t$. So we can derive the following theorem as a corollary.

\thmparens{}{K\H onig's theorem} Let $G$ be a bipartite graph with maximum matching $M$ and minimum vertex cover $C$. Then $|M| = |C|$.\slug

\section 3.3. Graph connectivity

\vskip -\medskipamount

\section 3.3.1. Definitions and inequalities

A graph $G=(V,E)$ is {\it connected} if there is a path between any two vertices. Otherwise, we say $G$ is {\it disconnected}. A {\it component} of $G$ is a maximal connected subgraph. If $G$ is connected and removing a set $W$ of vertices or edges causes it to become disconnected, then we say $W$ {\it separates} $G$. We have special names for $W$ if $|W| = 1$. If such a $W\subseteq V$, we call it a {\it cut vertex}. If $W\subseteq E, |W| = 1$, we call $W$ a {\it bridge}.

A graph $G=(V,E)$ is {\it k-connected} if $|V|>k$ and no vertex set $W\subseteq V$ of size $|W|<k$ separates $G$. Note that the complete graph $K_n$ is $(n-1)$-connected and that if $G$ is $k$-connected then it is also $k'$-connected for any $k'\leq k$. If $|V|\geq 2$ and no edge set $W\subseteq E$ of size $|W|<k$ separates $G$, then $G$ is {\it k-edge-connected}.

The {\it connectivity} of a graph $G$, denoted $\kappa(G)$, is the maximum $k\in \N$ such that $G$ is $k$-connected. Likewise, the {\it edge-connectivity} $\lambda(G)$ of a graph $G$ is the maximum $k$ such that $G$ is $k$-edge-connected. Note that if $G$ is not a complete graph, then both $\kappa(G)$ and $\lambda(G)$ are at most $|V|-2$.

\proclaim Lemma. Let $G=(V,E)$ be a graph. Then for all edges $uv\in E$, $\kappa(G\setminus\{uv\})\geq \kappa(G) - 1$.

\proof Remove an edge $uv$ from $G$. Now we find a set of vertices $W$, with $|W| = \kappa(G\setminus\{uv\})$, that disconnects $G$. So $V$ has been split into three disjoint subsets $L\cup W\cup R$. If $u$ and $v$ both lie in $L$ or both lie in $R$, then $\kappa(G) \leq |W| = \kappa(G\setminus\{uv\})$. If $u\in L$ and $v\in R$ (or vice-versa), then if there exists a vertex $w\neq u\in L$, then $W\cup \{u\}$ separates $G$, implying that $\kappa(G)\leq |W\cup \{u\}|\leq \kappa(G\setminus\{u,v\}) + 1$. The last case is if $L = \{u\}$ and $R=\{v\}$. Then $|V| = |W\cup\{u,v\}|\leq \kappa(G\setminus\{u,v\} + 2$, meaning that $\kappa(G)\leq |V|-1 \leq \kappa(G\setminus\{u,v\})+1$. \slug

\proclaim Lemma. Let $G=(V,E)$ be a graph with $|V|=n$. Then the following both hold:
\medskip
\item {1.} $\kappa(G) - 1 \leq \kappa(G-v)$ for all $v\in V$.
\smallskip
\item {2.}$\lambda(G) - 1 \leq \lambda(G-e)\leq \lambda(G)$ for all $e\in E$.
\medskip

\proof We prove each part separately.
\medskip
\item {1.} If $G = K_n$, then $G-v=K_{n-1}$, so $\kappa(G) - 1 = n-2 = \kappa(G-v)$ and we are done. So assume $G\neq K_n$. If $G\neq K_n$ and $G-v = K_{n-1}$, then there must exist some vertex $u\in V$ such that $uv\notin E$. Then removing the set $W = V\setminus\{u,v\}$ separates $G$, so $\kappa(G) \leq |W| = n-2 = \kappa(G-v)$. The last case is that $G\neq K_n$ and $G-v \neq K_{n-1}$. Then there exists a $W$ with $|W|=\kappa(G-v)$ such that $W\cup \{v\}$ separates $G$, which means that $\kappa(G) \leq |W\cup\{v\}| = \kappa(G-v)+1$.
\smallskip
\item {2.} There exists a set $W\subseteq E$ with $|W| = \lambda(G-e)$. So $e\notin W$ but $W\cup\{e\}$ separates $G$. So $\lambda(G) \leq |W\cup\{e\}| = \lambda(G-e) + 1$. For the second inequality, let $W$ be the set that separates $G$, so $|W|=\lambda(G)$. Then $W\setminus\{e\}$ separates $G-e$. So $\lambda(G-e) \leq |W\setminus\{e\}|\leq |W|=\lambda(G)$.\slug

To summarise, removing an edge from $G$ causes both $\kappa(G)$, $\lambda(G)$ to decrease by {\it at most} 1. Removing a vertex from $G$ can cause $\kappa(G)$ to decrease by at most 1 as well, but may cause $\lambda(G)$ to decrease a lot (possibly down to 0). The next lemma describes the relationship between vertex- and edge-connectivity.

\proclaim Lemma. Let $G=(V,E)$ be a graph and let $\delta(G)$ denote the minimum degree over all vertices in $G$. Then
$$ \kappa(G) \leq \lambda(G) \leq \delta(G).$$

\proof The inequality $\lambda(G) \leq \delta(G)$ is easy, because if $v$ is the vertex of minimum degree in $G$, then removing the $\delta(G)$ edges adjacent to $v$ separates $v$ from the rest of the graph.

To show that $\kappa(G) \leq \lambda(G)$, we use that $\kappa(G-e) \geq \kappa(G)-1$ for all $e\in E$. Suppose $\lambda(G) = n$. So removing a set of edges $\{e_1,\ldots,e_n\}$ causes $G$ to be disconnected. This means that $0=\kappa(G-\{e_1,\ldots,e_n\}) \geq \kappa(G) - n$, which means that $\kappa(G) \leq n = \lambda(G)$. \slug

\section 3.3.2. Connectivity and paths

For the following important theorem, we will need a new definition. We say two s-t-paths $P$ and $Q$ are {\it independent} if $V(P)\cap V(Q) = \{s,t\}$. The paths are {\it edge-disjoint} if $E(P)\cap E(Q) = \emptyset$.

\thmparens{M}{Menger}
Let $G$ be a graph. The following both hold:
\medskip
\item {1.} $G$ is $k$-connected if and only if there are $k$ independent paths between any two vertices $s$ and $t$.
\smallskip
\item {2.} $G$ is $k$-edge-connected if and only if there are $k$ edge-disjoint paths between any two vertices $s$ and $t$.
\medskip

\proof The ``if'' direction is left as an exercise for the student. To prove the ``only if'' direction, we use the max-flow min-cut theorem. We construct a directed graph $G'$ from $G$:
$$\eqalign{
    V(G') &= V(G)\cr
    E(G') &= \{uv, vu : uv\in E(G)\}
}$$
To prove the vertex version of the theorem, for every vertex $v\in V'\setminus\{s,t\}$, we set a vertex capacity $d(v) = 1$. To prove the edge version of the theorem, we set $c(e) = 1$ for all edges in $E(G')$. Now we consider the vertex and edge cases separately.
\medskip
\item {1.} Assume that $st\notin E(G')$. Any s-t-vertex cut of $G'$ separates $G$; since any s-t-path in $G'$ is also a t-s-path, the s-t-vertex cut also separates $t$ from $s$. Let $T$ denote a minimum-cardinality s-t-vertex-cut in $G'$ and we get that $d(T) \geq \kappa(G)$. There exists a 0-1 flow with $|f|\geq \kappa(G)$. So at each vertex, the number of incoming edges is at most 1 and the number of outgoing edges is at most 1 as well. This implies that there are $|f|$ independent pats from $s$ to $t$. But by the max-flow min-cut theorem, $|f| = d(T) \geq \kappa(G) = k$. If it so happens that $st\in E(G)$, then $(s, st, t)$ is an additional independent s-t-path to those in $G-st$ and we know that $\kappa(G-st)\geq \kappa(G) - 1$. Hence $G$ being $k$-connected implies that there are at least $k$ independent paths between any $s, t$.
\smallskip
\item {2.} In the case of edge capacities, any s-t-edge-cut separates $G$ as any t-s-path is also an s-t-path in G'. Let $S$ denote a minimum s-t-edge cut. Then $c(E)\geq \lambda(G)$. Again, there exists a 0-1 flow f with $|f|\geq \lambda(G)$, but this time it may be possible that a vertex has many incoming and outgoing edges. However, we know that the number of incoming edges that carry flow must equal the number of outgoing edges that carry flow at any vertex. Let $H=(V,F)$ be a subgraph of $G$, where $F = \{e\in E(G) : f(e) = 1\}$. While there exists an s-t-path $P$ in $H$, we can remove any all edges of $P$ from $H$ while maintaining flow conservation at every vertex $v\in V\setminus\{s,t\}$. We can repeat the algorithm as many times as there are edges coming out of $S$, so at least $|f|$ times. Hence $|f|\geq \lambda(G)\geq k$.\slug

\section 3.3.3. 2-connectivity and ear decompositions

An {\it ear decomposition} of a graph $G$ is a sequence $G_0, G_1, \ldots, G_k$ of subgraphs of $G$ such that $G_0$ is a cycle and for any $i\in \{1,\ldots,k\}$, the graph $G_i$ is obtained from $G_{i-1}$ by adding a path $P_i$ to $G_{i-1}$ such that $P_i$ shares exactly its endpoints with $G_{i-1}$. Each of these paths $P_i$ is called an {\it ear}.

\proclaim Theorem. A graph is 2-connected if and only if it has an ear decomposition.

\proof The only way for a graph to not be 2-connected is for it to have a cut vertex. So the ``if'' direction is easy, because if there is an ear decomposition, there is no cut vertex.

To prove the ``only if'' direction, suppose that $G$ is a 2-connected graph, and let $G_0$ be an arbitrary cycle in $G$. To get from any subgraph $G_{i-1}$ to $G_i$, we need to find an ear. Since $G$ is connected and $G_{i-1}\neq G$, there exists an edge $uv\in E(G)\setminus E(G_{i-1})$ such that $u\in V(G_{i-1})$. If $v\in V(G_{i-1})$, then $uv$ is an ear and we can set $P_i = uv$. Otherwise, there exists an edge $uw\in E(G_{i-1})$. This we know because $G_0\subseteq G_{i-1}$ so $|V(G_{i-1})| \geq 3$ and $G_{i-1}$ is connected.

By Menger's theorem, there exist at least 2 independent paths between $v$ and $w$ in $G$. So there exists some v-w-path $P$ such that $u\notin V(P)$. Let $u'$ be the first vertex of $P$ that is in $G_{i-1}$. Then we can append $u$ to the front of $P$ to get a path $(u, uv, v, \ldots, u')$, which is an ear.\slug

For a graph $G$, we define an {\it edge addition} is the operation of adding an edge between two vertices that were not connected in $G$. An {\it edge subdivision} is the operation of splitting an edge into two edges, with a new vertex in between them.

\proclaim Corollary. A graph is 2-connected if it can be obtained from $K_3$ by a sequence of edge additions and subdivisions.

This corollary is immediate from the observation that any starting cycle $G_0$ is a subdivision of $K_3$, and any ear $P_i$ is a subdivision of an edge addition.

\section 3.4. Counting spanning trees

A {\it spanning tree} $T$ of a graph $G$ is a subgraph that contains all vertices in $G$ and is a tree with edges from $G$. Let $T(G)$ denote the number of spanning trees of a graph $G$. For example, $T(K_3) = 3$. We want to find out the value of $T(G)$ for different classes of graphs.

In our investigation of $T(K_n)$, we will need to define a certain type of finite sequence. A Pr{\"u}fer code is a sequence $(p_1, \ldots, p_{n-2})\in \{1,\ldots,n\}^{n-2}$. It turns out that there is a bijection between Pr{\"u}fer codes and trees of length $n-2$ and trees on $n$ vertices. First we consider the algorithm for getting the Pr{\"u}fer code from a tree on $n$ vertices:

\algbegin Algorithm P (Build Pr{\"u}fer code). Given a tree with $n$ vertices $T$, this algorithm constructs a Pr{\"u}fer code.

\algstep P1. [Initialise.] Set $i\gets 1$, $T_0\gets T$.
\algstep P2. [Get next element.] Let $l_i$ denote the leaf of $T_{i-1}$ with smallest index. Set $p_i$ equal to the (single) neighbour of $l_i$ in $T_{i-1}$.
\algstep P3. [Remove a leaf.] Set $T_i\gets T_{i-1}-l_i$.
\algstep P4. [Done?] If $i=n-2$, we output the sequence $(p_1,\ldots, p_{n-2})$. 
\algstep P5. [Repeat.] Otherwise, set $i\gets i+1$ and return to step P2. \slug

There is a reverse algorithm that builds a tree from a Prufer code:

\algbegin Algorithm T (Build a tree). Given a Pr{\"u}fer code $(p_1,\ldots, p_{n-2})$, this algorithm outputs a tree with $n$ labelled vertices.

\algstep T1. [Calculate $l_i$.] For every $i\in \{1, \ldots, n\}$, let $l_i$ denote the number of times $i$ appears in the Pr{\"u}fer code, plus 1.
\algstep T2. [Initialise.] Let $T_0 \gets(\emptyset, \emptyset)$ to begin with. Set $i\gets 1$.
\algstep T3. [Add an edge.] Let $j$ be the smallest index with $l_j = 1$. Set $T_i \gets T_{i-1} + \{p_i, j\}$, where the addition operation denotes adding the edge as well as the vertices $p_i, j$, if one (or both) of them is not yet in the tree.
\algstep T4. [Update $l$.] Set $l_{p_i} \gets l_{p_i} - 1$ and $l_j \gets l_j - 1$.
\algstep T5. [Done?] If $i = n-2$, we have worked through the whole Pr{\"u}fer code. There are two remaining non-zero elements in $l$, call them $l_x$ and $l_y$. Output $T_i + \{l_x, l_y\}$.
\algstep T6. [Repeat.] Otherwise, set $i\gets i+1$ and return to step T3. \slug

Since Algorithms P and T are both deterministic and they are inverse to one another, this implies a bijection between trees on $n$ vertices and Pr{\"u}fer codes of length $n-2$. Hence we have the following formula.

\thmparens{C}{Cayley's formula} The number of spanning trees of a complete graph with $n$ vertices is
$$ T(K_n) = n^{n-2}.$$

\proof Since every possible edge between two vertices exists in $K_n$, the number of spanning trees of $K_n$ is exactly the number of trees with $n$ vertices. This is the same value as the number of Pr{\"u}fer codes of length $n-2$, which is the value $n^{n-2}$, since there are $n$ choices for each of $n-2$ elements.\slug

The following theorem gives a formula for the number of spanning trees of $K_n$, after a single edge is removed.

\proclaim Theorem. Let $e$ be an edge in $K_n$. Then
$$ T(K_n-e) = (n-2)n^{n-3}.$$

\proof The idea is to double-count the number of edges in all spanning trees of $K_n$. Each spanning tree has $n-1$ edges so the total number of edges is $(n-1)n^{n-2}$. Now let $k_e$ denote the number of spanning trees of $K_n$ containing $e$. By symmetry, this is the same for any choice of $e$ in $K_n$. There are ${n\choose 2}$ edges in $K_n$, each contained in $k_e$ trees. This implies that
    $$ {n\choose 2}k_e = (n-1)n^{n-2}.$$
    Solving for $k_e$, we obtain the value $k_e = 2n^{n-3}$. Therefore $T(K_n - e) = T(K_n)-k_e = n^{n-2} - 2n^{n-3} = (n-2)n^{n-3}$, which is what we wanted.\slug

\section 4. EXTREMAL THEORY

\vskip -\medskipamount

\section 4.1. Extremal graphs

This section deals with the general question: ``What is the extremal (maximum/minimum) number of objects, subject to restriction $R$?'' Some simple examples from graph theory include
\medskip
\item {$\bullet$} Question: What is the maximum number of edges in a graph with $n$ vertices? Answer: ${n\choose 2}$.
\smallskip
\item {$\bullet$} Question: What is the maximum number of edges in a graph with no cycles? Answer: $n-1$.
\medskip

A harder question is ``What is the maximum number of edges in a graph not containing $C_3$ as a subgraph?'' First let us consider a class of graphs that definitely do not contain $C_3$ as a subgraph: bipartite graphs (they do not contain any odd cycles). What is the maximum number of edges in a bipartite graph? Well, the complete bipartite graph $K_{a,b}$ has $ab$ edges, so we want to maximise the product $ab$ when $a+b =n$. Intuitively this is when $a$ and $b$ are as close to $n/2$ as possible. So we get that the number of edges in $K_{a,b}$ is at least $\lfloor n^2/4 \rfloor$, thus we have a lower bound for the number of edges in a graph not containing $C_3$. The following theorem proves that the bound is tight.

\proclaim Theorem. The maximum number of edges of any graph $G$ not containing $C_3$ as a subgraph is at most $\lfloor n^2/4 \rfloor$.

\proof We want to show that there exist $a,b\in \N$ such that $|E(G)|\leq |E(K_{a,b})| \leq \lfloor n^2/4 \rfloor$. To this end, we find disjoint subsets $A,B\subseteq V(G)$ such that $A\cup B = V(G)$. and let $H$ = $K_{a,b}$ on $A$ and $B$.

Then it suffices to show that for any $v\in V(G)$, $\deg_G(v) \leq \deg_H(v)$, since the number of edges in any graph equals the sum over degrees of vertices, divided by 2. So let $v_0$ be a vertex of maximum degree in $G$. Then we can set $B = N_G(v_0)$ and $A = V\setminus B$ Now for any $v\in A$, $\deg_H(v) = |B| = \deg_G(v_0) \geq \deg_G(v)$. Note that in $G$, no two vertices of $B$ are adjacent, since then there would be a cycle of length 3. Then for any $v\in B$, $N_G(v)\subseteq A$, hence $\deg_G(v)\leq |A| = \deg_H(v)$.\slug

Compared to ${n\choose 2}$, $\lfloor n^2/4 \rfloor$ is still quite a lot, but remember that disallowing any cycles at all, we have a much smaller bound of $n-1$. A natural question is ``How long do forbidden cycles have to be so that the maximum number of edges is much smaller than ${n\choose 2}$?'' As we will see from the following theorem, disallowing $C_4$ lowers this bound quite a bit.

\proclaim Theorem. The maximum number of edges of any graph $G$ not containing $C_4$ as a subgraph is at most $(n^{3/2} + n) / 2$.

\proof The idea is to double-count the size of the set $M$ of pairs $(\{u, u'\}, v)$, where $u\neq\ u'\neq v$ and $uv, u'v\in E$. Note that for any set $\{u, u'\}$, there is at most one $v\in V$ such that $(\{u, u'\}, v)\in M$, because otherwise we would have a cycle of length 4, which is forbidden. This implies that $|M|\leq {n\choose 2}$, which is the number of sets $\{u, u'\}$.

Another way to count $|M|$ is to consider the number of sets $\{u, u'\}$ that are contributed to by each $v\in V$. For every set $\{u, u'\}\subseteq N(v)$ we get a pair $(\{u, u'\}, v)\in M$, so $v$ contributes ${\deg(v)\choose 2}$ elements to $M$. For ease of notation, let us number the vertices $V = \{1, \ldots, n\}$ and let $d_i$ denote $\deg(i)$. Then
$$ |M| = \sum_{i=1}^n {d_i\choose 2}\leq {n\choose 2} .$$
To get a bound on the number of edges, we will relate $(\sum_{i=1}^n d_i)/2$ to $|M|$. We know that ${n\choose 2}\leq n^2/2$. Without loss of generality, we may assume that $d_i\geq 1$ for all $i\in V$, since adding an edge between a vertex $i$ with $d_i = 0$ to any other vertex $j$ increases the number of edges without introducing a $C_4$. Therefore,
$$ {d_i\choose 2} = {d_i(d_i -1)\over 2} \geq {1\over 2}(d_i - 1)^2 $$
for every $i\in V$. This implies that
$$ \sum_{i=1}^n (d_i - 1)^2 \leq n^2. $$
Now we use the famous {\it Cauchy-Schwarz inequality}, which applies to vectors $(x_1, \ldots, x_n)$, $(y_1, \ldots, y_n)\in \R^n$:
$$\sum_{i=1}^n x_iy_i \leq \sqrt{ \sum_{i=1}^n x_i^2 }\sqrt{ \sum_{i=1}^n y_i^2 }.$$
We apply the formula with $x_i = d_i - 1$ and $y_i = 1$ to obtain
$$\sum_{i=1}^n (d_i - 1) \cdot 1 \leq \sqrt{\sum_{i=1}^n (d_i-1)^2} \sqrt{\sum_{i=1}^n 1} \leq \sqrt{n^2}\sqrt{n} = n^{3/2}.$$
Because $\sum_{i=1}^n (d_i-1) = \sum_{i=1}^n d_i - n$, we get $\sum_{i=1}^n d_i \leq n^{3/2} + n$. So we can put a bound on the number of edges:
$$ |E| = {1\over 2}\sum_{i=1}^n d_i \leq {n^{3/2} + n\over 2} .\noskipslug$$

\section 4.2. Partially-ordered sets

A {\it partially-ordered set} or {\it poset} is a pair $({\cal L}, \subseteq)$ where $\subseteq$ is a partial order on the set ${\cal L} \subseteq 2^X$ over $X = \{1,\ldots , n\}$. A {\it chain} is a subset $\{A_1, \ldots, A_k\}$ of ${\cal L}$ such that $A_1\subseteq A_2\subseteq\cdots\subseteq A_k$. An {\it antichain} or {\it independent set system} is a subset $\{A_1, \ldots, A_k\}$ of ${\cal L}$ such that $A_i\not\subset A_j$ for all $i,j\in\{1,\ldots, k\}, i\neq j$. A chain is {\it maximal} if adding any set breaks the chain property.

The maximum length of a chain in $({\cal L}, \subseteq)$ is $|X| + 1$, where $X$ denotes the motherset. This is because $\emptyset$ can be part of the chain as well. Now consider the maximum length of an {\it antichain}. One way to get an antichain is to take all sets $A\in {\cal L}$ such that $|A| = i$ for some $i$. This number is at least ${n\choose i} \leq {n\choose \lfloor n/2 \rfloor}$ (in the case that ${\cal L} = 2^X$). The following theorem gives an upper bound:

\thmparens{S}{Sperner} Any antichain of a poset of $X = \{1, \ldots, n\}$ has size at most $n\choose \lfloor n/2 \rfloor$

\proof We will work with ${\cal L} = 2^X$. Let an antichain $M\subset 2^X$ be given. Our goal is to bound the cardinality of $M$. The idea is to double count the number of pairs $(R,A)$ where $A\in M$ and $R$ is a maximal chain containing $A$. Notice that a maximal chain contains exactly one set of size $i$ for $i\in \{0,\ldots, n\}$:
$$\emptyset \subseteq \{x_1\} \subseteq \{x_1, x_2\} \subseteq \cdots \subseteq \{x_1, x_2, \ldots, x_n\}.$$
where $x_1\ldots x_n$ are elements of $X$ written in some order. This implies that the number of maximal chains is $n!$ as every possible ordering defines a chain. By observation, any $R$ contains at most one $A\in M$, so the number of pairs $(R,A)$ is at most $n!$.

The other way we will count the pairs is to ask how many maximal chains actually contain a set $A\in M$. If $R$ is made up of elements $x_1,\ldots,x_n$ as shown above, then $A\in R$ if and only if $A = \{x_1,\ldots,x_k\}$ for some $k$. Hence to form $R$ we first introduce $x_1,\ldots,x_k$ in $k!$ ways, then $x_{k+1},\ldots,x_n$ in $(n-k)!$ ways, so the number of pairs $(R,A)$ is
$$ \sum_{A\in M} |A|! (n-|A|)! \leq n!$$
Dividing by $n!$ on both sides of the equation we get that
$$\sum_{A\in M} {|A|! (n-|A|)!\over n!} = \sum_{A\in M} {n\choose |A|}^{-1} \leq 1.$$
We know that ${n\choose |A|}\leq {n\choose \lfloor n/2 \rfloor}$, hence we may replace the reciprocal of the former with the reciprocal of the latter to get
$$1\geq \sum_{A\in M} {n\choose \lfloor n/2 \rfloor}^{-1} = {|M|\over {n\choose \lfloor n/2\rfloor}},$$
which means that $|M| \leq {n\choose \lfloor n/2 \rfloor}$.\slug

\section 5. FINITE PROJECTIVE PLANES

Let $X$ be a finite point set and ${\cal L}\subseteq 2^X$ be a set of subsets of $X$. $(X, {\cal L})$ is called a {\it finite projective plane} if the following properties hold:
\medskip
\item {P0)} There exists a set $F\subseteq X$ such that $|F| = 4$ and for all $L\in {\cal L}$, $|F\cap L|\leq 2$.
\smallskip
\item {P1)} For all $L_1,L_2\in {\cal L}$ with $L_1\neq L_2$, we have $|L_1\cap L_2| = 1$. 
\smallskip
\item {P2)} For all $x_1, x_2\in X$ with $x_1\neq x_2$, there is {\it exactly one} $L\in {\cal L}$ such that $x_1,x_2\in L$ and we may denote it $L = \overline{x_1x_2}$.
\medskip
We call any $x\in X$ a {\it point} and any $L\in {\cal L}$ a {\it line}. Finite projective planes have very interesting ``symmetrical'' properties.

\proclaim Lemma. Let $(X,{\cal L})$ be a finite projective plane. Then for all $L, L'\in {\cal L}$, there exists an $z\in L$ such that $x\notin L\cup L'$

\proof By (P0), there exists a set $F\subseteq X$ with $|F| = 4$ and $|F\cap L|\leq 2$ for a $|F\cap L'|\leq 2$. If $F$ is a proper subset of $L\cup L'$, then we're done, because $z\in F$. So assume $F\subset L\cup L'$: say $F = \{a, b,c,d\}$, $F\cap L = \{a, b\}$, and $F\cap L' = \{c, d\}$. Then by (P2), there exist lines $L_1 = \overline{ac}$ and $L_2 = \overline{bd}$; and by (P1), there exists a point $z\in L_1\cap L_2$. We claim that $z\notin L\cup L'$.

Suppose, towards a contradiction, that $z\in L$. Then since, by (P1), $|L\cap L_1| = 1$ and $z\in L_1$, we have that $z = a$. This implies that $L_2$ contains $a,b,d$. So $|F\cap L_2| \geq 3$, a contradiction to (P0). An analogous argument can be made for $L'$ so we are done. \slug

\proclaim Lemma. Let $(X,{\cal L})$ be a finite projective plane. Then for all $L,L'\in {\cal L}$, $|L| = |L'|$.

\proof The idea is to find a bijection $\phi : L \rightarrow L'$. By the preceding lemma, there exists some $z\notin L\cup L'$, so for any $x\in L$, we let $\phi(x)$ be the point in $L'\cap \overline{zx}$. The function $\phi$ is well-defined, since $\overline{zx}$ exists by (P2), and $|L' \cap \overline{zx}| = 1$ by (P1).

Now we prove that $\phi$ is a bijection. Let $y\in L'$ be given. It suffices to show that $|\phi^{-1}(y)| = 1$. Let $x\in L\cap \overline{yz}$. Then the set $\{x,z\}\subseteq \overline{yz}\cap\overline{xz}$, i.e.\  $|\overline{yz}\cap\overline{xz}|\geq 2$, so by (P1), $\overline{yz} = \overline{xz}$. This means that $\phi(x) = y$ and $x$ is unique by (P1), so $|\phi^{-1}(y)| = 1$.\slug

The {\it order} of a finite projective plane $(X, {\cal L})$ is $|L| - 1$ for any $L\in X$.

\proclaim Theorem. Let $(X,{\cal L})$ be a finite projective plane of order $n$. Then the following all hold:
\medskip
\item {1.} Exactly $n+1$ lines pass through any point $x\in X$.
\smallskip
\item {2.} $|X| = n^2 + n + 1$.
\smallskip
\item {3.} $|{\cal L}| = n^2 + n + 1$.
\medskip

\proof First we show that for any $x\in X$, there exists some line $L\in {\cal L}$ such that $x\notin L$. By (P0), there exists some $F = \{a,b,c,d\}$ and without loss of generality we will assume that $x\notin\{a,b,c\}$. So $\overline{ab} \cap F = \{a,b\}$ and $\overline{ac}\cap F = \{a, c\}$. Now if $x = d$, then $x\notin \overline{ab}$ and we're done. If $x\in \overline{ab}$, then $x\notin\overline{ac}$ and if $x\in \overline{ac}$, then $x\notin\overline{ab}$, since $|\overline{ab} \cap\overline{ac}| = 1$.

Let $x\in X$ be given and let $L$ be a line in ${\cal L}$ such that $x\notin L$. Now we may prove each part of the theorem.
\medskip
\item {1.} For all $y\in L$, there exists a line $\overline{xy}$ passing through $x$. There are $n+1$ points $y\in L$, so there are at least $n+1$ lines passing through $x$. Now for all $L'\in{\cal L}$ such that $x\in L'$, $|L\cap L'| = 1$ by (P1), so $L'$ was already counted above. So there are exactly $n+1$ lines through $x$.
\smallskip
\item {2.} Let $L_1 = \overline{xx_i}$ where $x_i\in L, i\in\{1,\ldots, n+1\}$. By (P1), $L_i\cap L_j = \{x\}$ for all $i\neq j$. As $|L_i\setminus\{x\}| = n$ we have
    $$|X|\geq |\bigcup_{i=1}^{n+1} L_i| = (n+1)n + 1 = n^2 + n + 1,$$
    so $|X|\geq n^2 + n + 1$.
    Now we show that for all $p\in X$, there exists an $i\in \{1,\ldots,n+1\}$ such that $p\in L_i$. If $p=x$. this is clearly true, so assume $p\neq x$. Then $\overline{px}\cap L = \{x_i\}$ for some $i$ by (P1). So $\overline{px} = L_i$ for some $i$ and we already counted $p$ above. So $|X| = n^2 + n + 1$.
\smallskip
\item {3.} This follows from item 2 and duality, which we will introduce next.\slug

The {\it incidence graph} of a finite projective plane $(X,{\cal L})$, is a bipartite graph $G = (V, E)$ with $V = X\cup {\cal L}$ and $E = \{xL : if x\in L\}$. The concept of duality involves switching the roles of $X$ and ${\cal L}$ in this bipartite graph, interpreting $X$ as lines and ${\cal L}$ as points.

Formally, the {\it dual} of a finite projective plane $(X,{\cal L})$ is the set system $({\cal L}, \Lambda)$ where $\Lambda\in 2^{\cal L}$ contains an element $\{L\in {\cal L} : x\in L\}$ for every $x\in X$.

\proclaim Lemma. The dual $({\cal L}, \Lambda)$ of a finite projective plane $(X,{\cal L})$ is also a finite projective plane.

\proof We show that $({\cal L}, \Lambda)$ satisfies each of (P0), (P1), and (P2).
\medskip
\item {(P1)} We need to find lines $L_1, L_2, L_3, L_4$ such that any for any $\lambda\in \Lambda$, with $\lambda = \{L\in {\cal L} : x\in L\}$ for some $x\in X$, the point $x$ is contained in at most two of $L_1, L_2, L_3, L_4$. Let $F = \{a,b,c,d\}$ be the set given by applying (P0) to $(X,{\cal L})$. Let $L_1 = \overline{ab}, L_2 = \overline{cd}, L_3 = \overline{ad}, L_4 = \overline{bc}$. Suppose, towards a contradiction, that $x$ is in three of these lines. Without loss of generality, suppose $x\in L_1\cap L_2\cap L_3$. By (P1) for $(X,{\cal L})$, $|L_i\cap L_j|$ for all $i\neq j$. Since $x\in L_1\cap L_3$, we have that $x = a$. But since $x\in L_2\cap L_3$, we have that $x = d$. This is a contradiction, since $a\neq d$. The same is true for any other triple of these four lines.
\smallskip
\item {(P2)} Let $\lambda_1,\lambda_2\in \Lambda, \lambda_1\neq\lambda_2$ be given, i.e.\  $\lambda_1 = \{ L\in {\cal L} : x_1\in L\}$ and $\lambda_2 = \{L\in {\cal L} : x_2\in L\}$ for some points $x_1\neq x_2$. We need that $|\lambda_1\cap\lambda_2| = 1$, but this follows from (P2) for $(X,{\cal L})$, since $x_1$ and $x_2$ intersect at exactly one line.
\smallskip
\item {(P3)} Let $L_1, L_2\in {\cal L}$ with $L_1\neq L_2$. It suffices to show that there exists a unique $\lambda \in \Lambda$ such that $L_1\in \lambda$ and $L_2\in \lambda$. But since $\lambda = \{L\in {\cal L} : x\in L\}$ for some $x$, this is exactly (P1) for $(X,{\cal L})$.\slug

\proclaim Corollary. The order of the dual finite projective plane is the same as the order of the original.

\proof For any $\lambda \in \Lambda$ of the dual, $|\{L\in {\cal L} : x\in L\}| = n+1$. This implies that the order or $({\cal L}, \Lambda)$ is $n$.\slug

One may wonder whether there is a finite projective plane of any order. The answer is no. There are no finite projective planes of order 1, 6, or 10. There are finite projective planes of order 3, 4, 5, 7, 8, 9, and 11. Whether there is a finite projective plane of order 12 is an open problem. The existence of finite projective planes of some higer orders is known though, as stated by the following theorem.

\proclaim Theorem. A finite projective plane exists of order $n$ if $n$ is a prime power.

We will not present the proof, which follows from ${\bf F}_n$ being a field.\slug

A corollary of this theorem is that there are infinitely many finite projective planes. Recall that a graph on $n$ vertices without $C_4$ as its subgraph has at most $(n^{3/2} + n)/2$ edges. Then a corollary of the infinity of finite projective planes is given below.

\proclaim Corollary. For infinitely many values of $n$ there is a graph on $n$ vertices without $C_4$ as a subgraph that contains at least $(n/2)^{3/2}$ edges.

\proof Consider the incidence graph of the finite projective plane $(X,{\cal L})$. The number of vertices $n = |X| + |{\cal L}| = 2(n^2 + n + 1)$. Then for all $L\in {\cal L}$, since $|L|=m+1$, the degree of $L$ in $G$ is $m+1$ as well. This means that $|E| = (m+1)|{\cal L}| = (m+1)(m^2+m+1) \geq (m^2+m+1)^{3/2} = (n/2)^{3/2}$. So the incidence graph has at least $(n/2)^{3/2}$ edges, so we just need to show that it does not have $C_4$ as a subgraph. This is easy because a $C_4$ in the bipartite incidence graph implies that the intersection of two lines $L\cap L'$ contains two points $x,x'$, which contradicts (P1).\slug

\section 6. LATIN SQUARES

A {\it Latin square} of order $n$ is a matrix $A \in \{1,\ldots, n\}^{n\times n}$ such that for any row $r$, $A_{ri} \neq A_{rj}$ if $i\neq j$ and for any column $c$, $A_{ic}\neq A_{jc}$ if $i\neq j$. Two Latin squares $A$ and $B$ are {\it orthogonal} if $(A_{ij}, B_{ij}) \neq (A_{xy}, B_{xy})$ whenever $i\neq x$ or $j\neq y$. Note that the number of ordered pairs of $\{1,\ldots,n\}$ is $n^2$ and there are only $n^2$ cells, so if all the pairs are different, each pair appears exactly once.

\proclaim Theorem. Let $M$ be a set of pairwise orthogonal Latin squares of order $n$. Then $|M|\leq n-1$.

\proof Let $A,B$ be orthogonal Latin squares of order $n$ and let $\pi$ be some permutation of $\{1,\ldots,n\}$. Consider $A'$ where $A'_{ij} = \pi(A_{ij})$. Note that $(A'_{ij}, B_{ij}) = (A'_{xy}, B_{xy})$ if and only if $(A'_{ij}, B_{ij}) = (A'_{xy}, B_{xy})$. So $A'$ and $B$ are also orthogonal.

So let $M = \{A_1, \ldots, A_t\}$ and for each $A_i\in M$, permute the entries such that the first row of the resulting Latin square $A'_i$ is $(1,2,3,\ldots,n)$. By our earlier observation, $A'_1,\ldots,A'_k$ are pairwise orthogonal. Now we zoom in on the second row, first column if $A'_i$ for some $i$. Call this cell $k$. First we notice that $k\neq 1$, since the first element of the first row is 1. Then since comparing any two $A'_i$ pairwise, we get all the pairs $(1,1), (2,2), \ldots, (n,n)$ in the first row, each of $2,3,\ldots,n$ can appear in the second row, first column of at most one $A'_i$. So $t = |M|\leq n-1$.\slug

The following theorem relates orthogonal Latin squares to finite projective planes.

\proclaim Theorem.
    For $n\geq 2$, a finite projective plane of order $n$ exists if and only if there exists a set of $n-1$ pairwise orthogonal Latin squares of order $n$.

\section 7. RAMSEY THEORY

Ramsey theory deals with the general question: ``How big must a mathematical structure have to be such that some property holds?"

\section 7.1. Independent sets and cliques

In a graph $G=(V,E)$, an {\it independent set} is a set $I\subseteq V$ such that $uv\notin E$ for all $u,v\in I$. A {\it clique} is a set $K\subseteq V$ such that $uv\in E$ for all $u,v\in K, u\neq v$. The {\it independent set number} $\alpha(G)$ is the maximum size of any independent set and the {\it clique number} $\omega(G)$ is the maximum size of any clique in $G$.

\proclaim Theorem. Let $G=(V,E)$ be a graph. If $|V|\geq {k-l-2\choose k-1} = {k+l-2\choose l-1}$, then $\alpha(G)\geq l$ or $\omega(G)\geq k$.

\proof By induction on $k+l$. In the base case $k = 1$ or $l=1$, all the theorem says is that there is an independent set or clique of size 4. For the inductive step, let $k,l\geq 2$ and assume that the claim holds for $k,l-1$ or $k-1,l$. By Pascal's formula,
$${k+l-2\choose k-1} = {k+l-3\choose k-1} + {k+l-3\choose k-2}.$$
Call the left-hand side $n$ and from the right-hand side, call the first summand $n_1$ and the second summand $n_2$. Now pick any vertex $u\in V$. Let $B = N(u)$ and $A = V\setminus(B\cup \{u\})$. First note that it is impossible that both $|A| < n_1$ and $|B|<n_2$ since that would imply that
$$n = 1 + |A| + |B| \leq 1 + (n_1-1) + (n_2=1) = n-1,$$
a contradiction. So either $|A|\geq n_1$ or $|B|\geq n_2$.

If $|A|\geq n_1$, by the induction hypothesis $\omega(G[A])\geq k$ or $\alpha(G[A])\geq l-1$. Then a clique in $G[A]$ is also a clique in $G$, for the first case; and for the second case, adding the vertex $u$ to an independent set in $G[A]$ gives an independent set in $G$, so $\alpha(G)\geq l$.

If it happens instead that $|B|\geq n_2$, then by induction we have either $\omega(G[B]) \geq k-1$ or $\alpha(G)\geq l$. In the first of these cases, an independent set in $G[B]$ is also an independent set in $G$, so $\alpha(G)\geq l$; and in the second case, adding $u$ to a clique in $G[B]$ gives a clique in $G$, so $\omega(G) \geq K$. \slug

For each $k,l\in \N$, the {\it Ramsey number} $r(k,l)$ is the minimum $n$ such that for any graph $G$ with $n$ vertices, $\omega(G)\geq k$ or $\alpha(G)\geq l$.

\proclaim Corollary. $$r(k,l) \leq {k+l-1\choose k-1} = {k+l-2\choose l-1}.$$

Very little is known about Ramsey numbers. By observation, we have that $r(k,1) = r(1,l) = 1$; $r(2,l) = l$ and $r(k,2) = k$. For $k=l$, we only know that $r(3,3) = 6$ and $r(4,4) = 18$. The exact values of $r(n,n)$ for $n\geq 5$ are unknown. The above corollary gave an upper bound and the next theorem gives a lower bound.

\proclaim Theorem. For $k\geq 3$, $r(k,k)>2^{k/2}$.

\proof First we introduce some basic notions from probability that we will need. Let $\Omega$ be a probability space and let $X$ be an event. Then $\Pr[X] = |X|/|\Omega| < 1$ if and only if $|X| < |\Omega|$, which would imply that $\neg X \neq \emptyset$.

Now we create a random graph $G$ by starting with a set of $n$ vertices and adding each edge with probablity $1/2$ uniformly at random. We're trying to prove that if $n\leq 2^{k/2}$ there is a $G$ with $|V| = n$ such that $\alpha(G) < k$ and $\omega(G) < k$, so it suffices to show that
$$\Pr[\alpha(G) \geq k\ \hbox{or}\ \omega(G)\geq k : n\leq k^{k/2}] < 1.$$
Let $A$ be a subset of vertices with $|A|=k$. Let $K_A$ denote the proposition ``A forms a clique'' and let $I_A$ denote the proposition ``A forms an independent set". Then let $X_A = K_A \cup I_A$. $\Pr[K_A] = (1/2)^{k\choose 2} = \Pr[I_A]$ and since $K_A \cap I_A = \emptyset$,
$$X_A = 2\cdot\left({1\over 2}\right)^{k\choose 2} = 2^{1-{k\choose 2}}.$$
Let $\Pr(Y)$ denote the probability that $\alpha(G) \geq k$ or $\omega(G) \geq k$. This is equal to the probablilty that $X_A$ holds for some $A\subseteq V$ with $|A|\geq k$. So
$$\Pr[Y] \leq \sum_{A\subseteq V, |A|\geq k} 2^{1-{k\choose 2}} = {n\choose 2}2^{1-{k\choose 2}} \leq {n^k\over k!}2^{1-{k\choose 2}},$$
and we can further derive the strict inequality
$${n^k\over k!}2^{1-{k\choose 2}} < {n^k\over {2^{k/2+1}}}2^{1-(k^2/2-k/2)} = \left({n\over 2^{k/2}}\right)^k,$$
which is less than $1$, since $n\leq 2^{k/2}$. \slug

\section 7.2. Increasing/decreasing subsequences

Given a finite sequence $S = (x_1, x_2, \ldots, x_n)$ of numbers, an {\it increasing} ({\it decreasing}) {\it subsequence} of length $k$ is a sequence $(x_{i_1}, x_{i_2}, \ldots, x_{i_n})$ such that for all $j,l$ with $1\leq j < l\leq k$, $i_j < i_l$ and $x_{i_j} \leq x_{i_l}$ ($x_{i_j} \geq x_{i_l}$).

\thmparens{E}{Erd\H os-Szekeres} For any finite sequence $S$ of at least $(r+1)(s+1) + 1$ numbers, there is an increasing subsequence of length $r$ or a decreasing subsequence of length $s$.

\proof Suppose, towards a contradiction, that every increasing/decreasing subsequence has length at most $r-1$/$s-1$ respectively. Then we label each $x_i$ of $S$ with a pair $(a_i, b_i)$ where $a_i$/$b_i$ is the length of the longest increasing/decreasing subsequence ending in $x_i$. (Note that each of the numbers $x_i, x_j$ have different labels.) Without loss of generality, assume $i<j$. If $x_i\leq x_j$, then $a_i < a_j$, as any increasing subsequence ending in $x_i$ can be extended by $x_j$. Likewise, if $x_j\leq x_i$, then $a_j < a_i$, as any increasing subsequence ending in $x_j$ can be extended by $x_i$.

By our assumption, $a_i\leq r-1$ and $b_i\leq s-1$ for all $i$. So the number of labels $(a_i, b_i)\leq (r-1)(s-1)$ which implies that the length of the sequence $S$ is at most $(r-1)(s-1)$, a contradiction.\slug

The given bound is tight; there are sequences with length $(r-1)(s-1)$ numbers that neither have an increasing subsequence of length $r$, nor have a decreasing subsequence of length $s$. To construct one, take a grid of $(r-1)(s-1)$ points in the plane and rotate it {\it very} slightly counterclockwise such that no two points are on the same horizontal or vertical line. Then we have a set of points $(x_i, y_i),\ldots(x_n,y_n)$ where $n=(r-1)(s-1)$. Now order the $y$-coordinates {\it by the value of the $x$-coordinates}. Any increasing subsequence can take at most one element from each column, so at most $r-1$, and any decreasing subsequence can take at most one element from each row, so at most $s-1$.

\section 8. ERROR-CORRECTING CODES

Suppose we have two people who need to communicate over a channel that is ``noisy'' in the sense that errors may be introduced into the message. To detect and solve this issue, checksums may be used. Suppose we have an alphabet $\Sigma = \{0,\ldots,q-1\}$ where $q$ is prime. Let $x$ denote a message $x\in \Sigma^k$ for some positive integer $k$. Then the {\it checksum} $c$ of $x$ is $\sum_{i=1}^k x_i \pmod{q}$. Then the message becomes $y\in {\bf F}_q^{k+1}$ where
$$ y_i = \cases{
    x_i & for $i\leq k$ \cr
    c   & for $i = k+1$
}$$
Since $q$ is prime, for any $x,y\in {\bf F}_q^kk$, if $x$ and $y$ differ at {\it exactly one} place, then their checksums will differ. But what happens if more errors occur? Also, is it possible for the receiver to determine where the error occured without asking the sender to resend the message?

To formalise the problem further, we introduce some more definitions. The {\it block length} $n$ is the length of the encoded message. Suppose the original (non-encoded) message has length $k$. Then the {\it encoding function} $E : \Sigma^k \rightarrow \Sigma^n $ is applied by the sender before transmission and the {\it decoding function} $D : \Sigma^n \rightarrow \Sigma^k$ is applied by the receiver upon the message's arrival. The {\it code} $C\subseteq \Sigma^n$ is the image of $E$. A code is {\it binary} if $\Sigma = \{0,1\}$. Checksums work by increasing the distance between two messages $x,y\in \Sigma^k$ if $x\neq y$.

The {\it Hamming distance} between two messages $x,y\in \Sigma^m$ is given by
$$ \Delta(x,y) = \sum_{i=1}^m [x_i\neq y_i].$$
For a code $C\subseteq \Sigma^n$, the {\it minimum distance} of $C$ is
$$ \Delta(C) = \min_{x,y\in C} \{\Delta(x,y)\}.$$
An {\it $(n,k,d)_q$-code} is a code $C$ such that
\medskip
\item {i)} $C\subseteq \Sigma^n$;
\smallskip
\item {ii)} $k = \log_q |C|$;
\smallskip
\item {iii)} and $d = \Delta(C)$.
\medskip
Note that $|C|$ is the size of the code and $k$ need not be an integer. A checksum is a $(n, n-1, 2)_q$-code: Suppose that $x\in C$. Let $y$ be such that $\Delta(x,y) = 1$. Is it possible that $y\in C$? No, because for all $x,y\in C$, if $x\neq y$ then $\Delta(x,y)\geq 2$.

If, for a code $C$, we can be sure that at most $r$ errors occur during transmission, then:
\medskip
\item {i)} If $\Delta(C) \geq r+1$, then the error can be {\it detected} because for any $x\in C$, if we have $y$ that differs from $x$ at not less than $r+1$ places, i.e.\  $\Delta(x,y)\leq r$, then $y\notin C$.
\smallskip
\item {ii)} If $r\leq \lfloor (\Delta(C) - 1)/2 \rfloor$, then the error can be {\it corrected}. Construct a graph with $V = \Sigma^n$ with an edge between $x,y\in \Sigma^n$ if $\Delta(x,y) = 1$. Now if we have $y\in \Sigma^n$ which we know is a garbled message, we can find the closest valid message in the graph. If the error is not too large, then the closest valid message will have been the intended one.
\medskip

The general aim is to construct codes with small $n$ and large $\Delta(C)$. For any finite projective plane $(X,{\cal L})$ of order $m$, we can construct a $(m^2+m+1, \log_2(m^2+m+1), 2m)_2$-code by associating every letter in the alphabet with a characteristic vector of $L\in {\cal L}$. For example, the letter associated with the line $\{1,5,6\}$ in the Fano plane will be coded as $1000110$. Since for any $L_1, L_2\in {\cal L}, L_1\neq L_2$, $|L_1\cap L_2| = 1$ and each of $L_1, L_2$ have $m+1$ points, $\Delta(y_{L_1}, y_{L_2}) = 2m$, where $y_{L_i}$ is the letter associated with the line $L_i$.

A {\it linear} code $C\subseteq {\bf F}_q^n$ is a linear subspace of ${\bf F}_q$, i.e.\  for all $x,y\in C$, $\alpha\in {\bf F}_q$, $\alpha x+y\in C$. There exist $k$ independent vectors $x_1,\ldots, x_k\in {\bf F}_q^n$ such that
$$C = \{\sum_{i=1}^k \alpha_ix_i : \alpha_i,\ldots \alpha_n\in {\bf F}_q\}.$$
More succinctly, we can express $C$ using the generator matrix $G\in {\bf F}_q^{n\times k}$. The matrix $G = \matrix{ x_1 & \cdots & x_k }$ and $C = \{G_\alpha : \alpha \in {\bf F}_q^k\}$. (Multiplying any message by $G$ on the left encodes it.)

Alternatively, represent $C$ by the {\it null space}. Recall that for any $k$-dimensional linear subspace $C$, there is an $(n-k)$-dimensional linear subspace $C^\perp$ such that for all $x\in C, y\in C^\perp$, $x^Ty = 0$. Using the generator matrix $H\in {\bf F}_k^{n\times (n-k)}$ of $C^\perp$, express $C = \{ x : x^TH = 0\}$. Then $H$ is called the {\it parity check} matrix of $C$. In a sense, linear codes are simple because we can detect errors simply by computing $x^TH$.

The {\it support} of a vector $x$ is the set $\{i : x_i \neq 0\}$ and the {\it Hamming weight} of $x$ is $\hbox{wt} x = |\{i : x_i \neq 0\}|$.

\proclaim Theorem. For a linear code $C$,
$$ \Delta(C) = \min\{\hbox{wt} x : x\in C\ \hbox{and}\ x\neq 0\}.$$

\proof For ease of notation, let $d$ denote $\min\{\hbox{wt} x : x\in C\ \hbox{and}\ x\neq 0\}$. First we show that $d\leq \Delta(C)$. Let $y,z\in C$ be such that $\Delta(y,z)$ = $\Delta(C)$. Since $C$ is linear, $x = y-z \in C$. Note that $\Delta(y,z) = |\{i : x_i \neq 0\}|$, so $d \leq \hbox{wt} x = \Delta(C)$.

Now we show that $d\geq \Delta(C)$. Note that $0\in C$. Let $x\neq 0$ be such that $\hbox{wt} x = d$. Then $\Delta(C)\leq \Delta(0,x) = |\{i : x_i \neq 0\}| = \hbox{wt} x = d$.\slug

We can use this theorem to construct a linear code with $\Delta(C) \geq 3$. We find a parity check matrix $H$ such that $x^TH \neq 0$ for any non-zero $x$ with $\hbox{wt} x\in \{1,2\}$. For simplicity, we will work with $q=2$. If $\hbox{wt} x = 1$, then $X$ has exactly one 1-entry, say it is at position $i$. So $x^TH$ is the $i$-th row $H_i$ of $H$. We need that $H_i$ is non-zero for every $i$. If $\hbox{wt} x = 2$, then $x^T$ has exactly two non-zero elements, say they are at positions $i$ and $j$. Then $x^TH$ = $(H_i + H_j)^T$ hence we need $H_i\neq H_j$ for all $i\neq j$.

If all rows of $H$ are distinct and non-zero, then we get $\Delta(C) \geq 3$. The largest number of such rows if $H\in {\bf F}_2^{n\times l}$ is $2^l-1$ so $n = 2^l-1$ for any $l = n-k$. Hence we obtain an $(n, n-\log_2(n+1), 3)_2$-code called the {\it Hamming code}.

Let $x\in {\bf F}_q^n$. The {\it ball} around $x$ of radius $r$ is given by
$$B(x,r) = \{y\in {\bf F}_q^n : \Delta(x,y) \leq r\}.$$
The {\it volume} of such a ball is
$$\hbox{vol}(r,n) = |B(x,r)| = \sum_{i=0}^r {n\choose i} (q-1)^i.$$

\thmparens{S}{Hamming bound} If an $(n,k,d)_q$-code exists, then
$$q^k\cdot\hbox{vol}\left(\lfloor {d-1\over 2} \rfloor, n\right) \leq q^n.$$

\proof Let $r = \lfloor (d-1)/2 \rfloor$ and consider $\bigcup_{x\in C} B(x,r)\subset \{1,\ldots,q\}^n$. By observation,
$$|\bigcup_{x\in C} B(x,r)| = \sum_{x\in C} \hbox{vol}(r,n) = q^k\cdot \hbox{vol}(r,n) \leq q_n.\noskipslug$$

Note that if $d = 3$ and $q = 2$ then the Hamming bound is $2^k(n+1)\leq 2^n$ and the Hamming code matches this bound with equality as $k=n-log_2(n+1)$. An $(n,k,d)_2$-code is called {\it perfect} if $q^k\cdot\hbox{vol}(\lfloor (d-1)/2 \rfloor, n) = q^n$. There are no codes with larger size $k$ than a $(n, n-d+1, d)_2$-code (even if we let $q>2$), as the following theorem shows.

\thmparens{B}{Singleton bound} If $C$ is an $(n,k,d)_q$-code, then $k\leq n-d+1$.

\proof Define a function $f:\Sigma^n\rightarrow\Sigma^{k-1}$ such that $f(x_1,\ldots, x_n) = (x_1,\dots, x_{k-1})$. As $|\Sigma^{k-1}| = q^{k-1}$ and $|C| = q^k > q^{k-1}$, there are $x,y\in C$, with $x\neq y$ and $f(x) = f(y)$. So $x$ and $y$ can only differ in the last $n-k+1$ entries. Since $d = \Delta(C) \leq \Delta(x,y) \leq n-k+1$, $k\leq n-d+1$.\slug

An $(n,k,d)_q$ code is called {\it maximum-distance separable} or MDS if $k = n-d+1$.

\end