### 2019-08-05 misc/study-notes/abstract-algebra.tex
Fixed typo in graph theory and added algebra

% Abstract Algebra Notes
% Based on lectures given in Fall 2003 by Prof. Benedict Gross
% Notes by Marcel Goh

\input fontmac
\input mathmac

% Useful functions

\def\Sym{\mathop{\rm Sym}\nolimits}

\leftrighttop{NOTES ON ABSTRACT ALGEBRA}{MARCEL GOH}

\maketitlenodate{Notes on Abstract Algebra\footnote{$^1$}{\eightpt Based on a series of lectures given at Harvard University in Fall 2003 by Prof. Benedict Gross.}}{by}{Marcel K.\ Goh}

\section 1. PRELIMINARY NOTIONS FROM LINEAR ALGEBRA

This set of notes assumes that the reader has had some exposure to linear algebra; the most crucial notions will be outlined in this section.

\section 1.1. Basic properties of square matrices

We shall mainly concern ourselves with matrices of the form
$$\pmatrix{
    a_{11} & \cdots & a_{1n} \cr
    \vdots & \ddots & \vdots \cr
    a_{n1} & \cdots & a_{nn}
}$$
where $n$ is some positive integer and each entry $a_{ij}\in \R$. We call the set of all $n\times n$ matrices with real entries $M_n(\R)$.

For two matrices $A$ and $B$, we may form their {\it sum} $A+B = (a_{ij} + b_{ij})$. (This notation means that at the $i$th row and $j$th column, the entry is the sum of $a_{ij}$ and $b_{ij}.)$. Given a real number $\alpha$, we obtain the scalar product $\alpha A  = (\alpha\cdot a_{ij})$ by multiplying every entry in $A$ by $\alpha$.

We can also {\it multiply} $n\times n$ matrices $A$ and $B$ with the following formula:
$$A\cdot B = (c_{ij})\quad\hbox{where } c_{ij} = \sum_{k=1}^n a_{ik}\cdot b_{kj}.\oldno 1$$
Since a matrix represents a linear transformation, multiplying matrices is like composing functions. If $S$ and $T$ are the transformations represented by the matrices $B$ and $A$, respectively, then the matrix product $A\cdot B$ can be thought of as the following composition of transformations:
$$ \matrix{ \R^n & \buildrel S\over \longrightarrow & \R^n & \buildrel T\over \longrightarrow & \R^n } $$
For matrices $A$ and $B$ the commutativity of addition
$$A + B = B + A \oldno 2$$
is valid, and for three matrices $A$, $B$, and $C$, the distributive law
$$A\cdot (B + C) = A\cdot B + A\cdot C\oldno 3$$
and the associativity of multiplication
$$A\cdot (B\cdot C) = (A\cdot B)\cdot C \oldno 4$$
may be proven to hold as well. In particular, ({\oldstyle 4}) is laborious to prove from the definition given in ({\oldstyle 1}), but easy to derive when reasoning about matrices as transformations.

Unlike addition, multiplication is not commutative: $A\cdot B$ does not equal $B\cdot A$ in general. To prove this, we simply note that
$$\pmatrix{ 0 & 1 \cr 0 & 0} \cdot\pmatrix{ 0 & 0 \cr 0 & 1 } = \pmatrix{ 0 & 1 \cr 0 & 0 } \hbox{ but } \pmatrix{ 0 & 0 \cr 0 & 1} \cdot\pmatrix{ 0 & 1 \cr 0 & 0 } = \pmatrix{ 0 & 0 \cr 0 & 0 }.$$
The matrix
$$I =
\pmatrix{
    1 & 0 & \cdots & 0 \cr
    0 & 1 & \cdots & 0 \cr
    \vdots & \vdots \ddots & 0 \cr
    0 0 0 & \cdots & 1
}$$
is called the {\it identity matrix} and has the property that $AI = IA = A$ for any matrix $A$. [When no ambiguity can arise, we often omit the $\cdot$ symbol when denoting a product.]

We say that a matrix $A$ is {\it invertible} if and only if there exists a matrix $B$ such that $AB = BA = I$; otherwise, it is called {\it singular.} Not all matrices are invertible. For example, the matrix ${\bf 0}$, all of whose entries are $0$, is not invertible in any dimension. The identity matrix is easily seen to be invertible (take $B = I$). Note that if an inverse matrix exists for a given matrix $A$, then it is unique, for if $AB = AB' = I$ and $B'A = BA = I$, then, multiplying the first identity by $B$ on the left, we arrive at $BAB = BAB'$, i.e.\ $B = B'$.

Since a $1\times 1$ matrix $(a)$ contains only one real number, it is invertible if and only if $a \neq 0$. A $2\times 2$ matrix $a\,b \choose c\,d$ is invertible if and only if $ad-bc = 0$, since for any such matrix,
$$\pmatrix { a & b \cr c & d } \pmatrix { d & -c \cr -c & a} = \pmatrix { ad-bc & 0 \cr 0 & ad-bc },$$
and we can multiply both sides by $1/(ad-bc)$ to obtain the identity on the right, provided that $ad-bc$ is nonzero.

In general, there exists a function $\det : M_n(\R) \rightarrow \R$ such that a matrix $A\in M_n(\R)$ is invertible if and only if $\det A\neq 0$. This determinant can be calculated as a sum over $n!$ terms; this formula will not be useful for our purposes.

\section 1.2. The general linear group

Let us now restrict our attention to a certain subset of square matrices, namely those whose determinant is nonzero. This set is called the {\it general linear group of degree $n$} and is denoted $GL_n(\R)$ when all matrix entries are real numbers. Remark that, since $(-1) + (1) = (0)$, this set is not closed under addition; scalar multiplication is also no longer a safe operation, since multiplying any matrix by $0$ results in a singular matrix.

In return for these two forfeited closure properties, we get closure under matrix multiplication.

\proclaim Proposition I. Suppose that two matrices $A$ and $B$ are invertible. Then their product $AB$ is also invertible.

\proof Consider $B^{-1}A^{-1}$ and the product $(B^{-1}A^{-1})(AB)$. By associativity of multiplication, this becomes $B^{-1}(A^{-1}A)B = B^{-1}IB = IB^{-1}B = II = I$. Alternatively, use the fact that $\det(AB) = \det(A)\det(B)$, which is nonzero because both $\det(A)$ and $\det(B)$ are nonzero.\slug

Thus the set $GL_n(\R)$, under the operation of matrix multiplication, has a multiplicative inverse $A^{_1}$ for every matrix $A$. It also contains an identity element $I$ and the multiplication operation is associative. These are the properties of a group.

\section 2. GROUPS

\vskip -\medskipamount

\section 2.1. Properties and basic examples

A {\it group} $G$ is a set on which is defined a rule of combination such that the product of two elements $g,h\in G$, denoted $g\cdot h$ or $gh$, is also in $G$. Furthermore, the following three properties must hold:
\medskip
\item {a)} Multiplication must be associative: for all $g,h,k\in G$, $(gh)k = g(hk)$.
\smallskip
\item {b)} There exists an identity element $e\in G$ such that $ge = eg = g$ for all $g\in G$. This element is also often denoted $1$.
\smallskip
\item {c)} For every element $g\in G$, there exists an inverse element $g^{-1}$ such that $gg^{-1} = g^{-1}g = e$.
\medskip
If $gh = hg$ for all $g,h\in G$, then we say the group $G$ is {\it commutative} or {\it abelian}.

The most familiar abelian group is the set of all integers, denoted $\Z$, under the operation of addition. It has $1$ as its identity, inverses $-a$ for every whole number $a$, and the commutative property; likewise, any vector space $V$ is also an abelian group under vector addition.

Group theory is intimately connected to the study of symmetries of an object. Let $T$ be a set and let $\Sym(T)$ denote the set of all bijections from $T$ to itself. This is called the {\it symmetric group} on $T$ as, under composition of functions, it obeys all the group axioms: It contains the identity transformation $e$ and every bijection $f$ has an inverse bijection $f^{-1}$. In some sense, the symmetric group is the most general group, because all other groups arise from adding restrictions to these bijections. For instance, $GL_n(\R) \subset \Sym(\R^n)$. If $T = \{1,\ldots,n\}$, then $\Sym(T)$ is called the symmetric group on $n$ letters and is denoted $S_n$. This group has $n!$ elements and is non-abelian when $n\geq 3$.

\section 2.2 Subgroups

\end