### 2020-10-09 about/vita.tex
Lots of changes

% Marcel Goh CV

% From cwebmac: Date and time
\def\today{\ifcase\month\or
  January\or February\or March\or April\or May\or June\or
  July\or August\or September\or October\or November\or December\fi
  \space\number\day, \number\year}
\newcount\twodigits
\def\hours{\twodigits=\time \divide\twodigits by 60 \printtwodigits
  \multiply\twodigits by-60 \advance\twodigits by\time :\printtwodigits}
\def\gobbleone1{}
\def\printtwodigits{\advance\twodigits100
  \expandafter\gobbleone\number\twodigits
  \advance\twodigits-100 }

\font\eightrm=cmr8
\font\elevensc=cmcsc11  % For section headers
\font\bigbold=cmb12  % Name
\font\tenb=cmb10
\font\mc=cmr9

\let\tenbf=\tenb

% For publication lists (adapted from Knuth's algorithms)
\newdimen\itemindent
\newif\ifitempar \itempartrue
\def\itemindentset#1{\setbox0\hbox{{\bf #1.\kern.25em}}\itemindent=\wd0\relax}
\def\pubbegin #1{\itemindentset{#11}} % when all have 1 digit
\def\ppubbegin #1{\itemindentset{#111}} % when 10 or more publications
\def\pubitem#1.{\ifitempar\smallskip\noindent\else\itempartrue
  \hskip-\parindent\fi
  \hbox to\itemindent{\bf\hfil #1.\kern.25em}%
  \hangindent=\itemindent\hangafter=1\ignorespaces}

\def\up#1{\leavevmode \raise.16ex\hbox{#1}}
\def\sectheader#1{{\bigskip\elevensc #1}\smallskip\hrule\medskip}
\def\leftright#1#2{{#1}\hfill{#2}\par}
\def\myname{{\tensc Marcel K. Goh}}
\def\CEE{{\mc C}}
\def\UNIX{{\mc UNIX}}
\def\thing{\item{$\bullet$}}

\parindent=0pt  % No indentation

\newcount\submittedcount
\submittedcount=1
\def\subitem{\pubitem S\the\submittedcount. \global\advance\submittedcount by 1}
\newcount\reportcount
\reportcount=1
\def\repitem{\pubitem R\the\reportcount. \global\advance\reportcount by 1}

\def\datethis{\eightrm Last updated \today\ at\ \hours.}

% =========== START ============ %

\footline{\phantom{\datethis}\hss\folio\hss\datethis}  % \phantom for centering

\leftright{\bigbold Marcel K.\ Goh}{3660 Rue Hutchison, Apt.\ 4}
\leftright{Email: marcel\_goh@yahoo.ca}{Montr\'eal, QC H2X 2H3, Canada}
\leftright{URL: marcelgoh.github.io}{+1\thinspace\up(825\up)\thinspace 440-0681}

\sectheader{Education}

\leftright{\bf McGill University}{\bf September 2017 -- April 2021}
\leftright{\sl BSc, Honours Mathematics and Computer Science, minor Linguistics}{\sl Montr\'eal, QC, Canada}
CGPA: 3.79/4.
\medbreak
\leftright{\bf Charles University}{\bf February -- June 2019}
\leftright{\sl Exchange semester, Faculty of Mathematics and Physics}{\sl Prague, Czech Republic}
Grade: 1 on all courses (highest score attainable).

\sectheader{Awards}

\leftright{\bf NSERC Undergraduate Student Research Award}{\bf May 2020}
\smallskip
\leftright{\bf Governor General's Academic Medal -- Bronze}{\bf June 2015}

\sectheader{Research Experience}

\leftright{\bf Probabilistic analysis of branching processes}{\bf May 2020 -- present}
\leftright{\sl Research group}{\sl McGill University}
\smallskip
Ongoing research on branching processes, headed by Luc Devroye. Studied estimation problems
on Galton-Watson trees. Attended informal weekly seminars over the summer, discussing
various topics related to branching
processes. (Gave a presentation at three of these: one on root estimation in Galton-Watson
trees, and two on generating functions and elementary analytic combinatorics.)

\medbreak
\leftright{\bf Sorting algorithms on manifolds}{\bf May -- August 2020}
\leftright{\sl Summer research project}{\sl McGill University}
\smallskip
Research project under the supervision of Michael Lipnowski.
Studied algorithms that investigate the topology of group actions on locally symmetric spaces.
Work involved writing of code in OCaml and PostScript that generates illustrations of certain quotient
spaces, in various models of hyperbolic geometry.
\medbreak

\leftright{\bf Interactive proofs}{\bf September -- December 2019}
\leftright{\sl Honours research project}{\sl McGill University}
\smallskip
Semester-long research project in the Computation and Logic Group, supervised
by Brigitte Pientka.
Proved a theorem in constructive logic concerning the formal verification
of the interactive proof assistant Harpoon and wrote OCaml code as part of
ongoing work on the functional programming language Beluga.
\medbreak

\leftright{\bf Bytecode compiler and interpreter}{\bf February -- June 2019}
\leftright{\sl Individual software project}{\sl Charles University}
\smallskip
Semester-long individual software project under the supervision of Adam Dingle.
Created OPythn, a bytecode compiler and virtual machine for a subset of Python.
OPythn includes support for lists, dictionaries, named and anonymous functions, objects, and classes.
It is implemented in OCaml using Ocamllex for lexing and Menhir for parsing.

\sectheader{Submitted Papers}

\pubbegin S
\subitem (with Anna M. Brandenberger and Luc Devroye)
Root estimation in Galton-Watson trees. {\sl arXiv preprint 2007.05681}.
Submitted July 2020.

\sectheader{Reports}

\pubbegin R
\repitem Typechecking proof scripts: making interactive proof assistants robust.
Honours project report, McGill University (Montr\'eal, Qu\'ebec, December 2019), 10 pp.

\filbreak

\sectheader{Work and Volunteer Experience}

\leftright{\bf Department of Mathematics and Statistics, McGill University}{\bf September 2019 -- present}
\leftright{\sl Student Grader}{\sl Montr\'eal, QC, Canada}
\smallskip
Grading of assignments in the following courses:
\begingroup\parindent=10pt
\smallskip
\thing Fall 2020: MATH 323 Probability, MATH 456 Honours Algebra 3
\smallskip
\thing Winter 2020: MATH 240 Discrete Structures
\smallskip
\thing Fall 2019: MATH 235 Algebra 1
\endgroup
\medbreak

\leftright{\bf Computer Science Undergraduate Society Helpdesk}{\bf September 2018 -- present}
\leftright{\sl Tutor}{\sl Montr\'eal, QC, Canada}
\smallskip
Hold twice-weekly office hours to tutor students in a variety of undergraduate courses.
Topics covered include elementary data structures and algorithms, command-line scripting,
and functional programming. Recipient of the Tomlinson Engagement Award for Mentoring.
\medbreak

\leftright{\bf McGill Student Emergency Response Team}{\bf October 2017 -- present}
\leftright{\sl First Responder}{\sl Montr\'eal, QC, Canada}
\smallskip
On call on a weekly basis to provide emergency medical care at campus residences overnight as well as at
university events such as frosh, sports games, and formals. Attend team training sessions twice a month to keep
first-aid skills up-to-date.
\medbreak

\leftright{\bf Society of Undergraduate Mathematics Students}{\bf May 2019 -- January 2020}
\leftright{\sl Vice President, Academic}{\sl Montr\'eal, QC, Canada}
\smallskip
Oversaw academic affairs within SUMS council and acted as liaison between the undergraduate community and
mathematics faculty. Duties included representing the student body at department meetings,
organising midterm and final review sessions,
and helping students with
academic concerns.
\medbreak

\leftright{\bf Bakir Contracting Corp.}{\bf May -- August 2018}
\leftright{\sl Painter}{\sl Edmonton, AB, Canada}
\smallskip
Exterior painting (siding, decks, fences, trim, etc.) for residential clients.
\medbreak

\leftright{\bf Singapore Armed Forces}{\bf August 2015 -- August 2017}
\leftright{\sl Infantryman}{\sl Singapore}
\smallskip
Held appointment of machine-gun team commander in the 3rd Battalion, Singapore Infantry Regiment.
Led a six-person team consisting of a medic, signaller, sensor, and two-machine gunners within a rifle platoon.
\medbreak

\leftright{\bf YoWoChAs Outdoor Education Centre}{\bf June -- August 2014}
\leftright{\sl Summer Camp Counsellor}{\sl Fallis, AB, Canada}
\smallskip
Led children aged 4--15 through various activities (e.g. archery, canoeing, zipline) at a sleepaway camp.
\medbreak

\sectheader{Skills}

{\bf Programming Languages}\par
\CEE, OCaml, Haskell, Standard ML, Scheme Lisp, PostScript, {\tt CWEB}, {\tt MIXAL}, Python.
\medbreak

{\bf Technologies}\par
\UNIX, Vim, \TeX, Git.
\medbreak

{\bf Languages}\par
Fluent: English, French. Proficient: Mandarin, Italian.
\medbreak

\sectheader{Interests}

Analysis and design of algorithms, analysis of random discrete structures, combinatorics, group theory,
programming language theory and design, compilers.

\sectheader{Other}

\parindent=10pt
\thing Scored 10/120 on the William Lowell Putnam Mathematical Competition in 2019
(the median score that year was 2/120).
\smallskip
\thing Recipient of 0x\$0.40 in Knuth reward cheques.
\medbreak

\filbreak

\bye