### 2022-03-26 misc/expo/additive.tex
It is 2022 omg

\input fontmac
\input mathmac

\def\FF{{\bf F}}
\def\TT{{\bf T}}
\def\bar{\overline}
\def\hat{\widehat}
\def\norm#1{|\!|#1|\!|}
\def\bignorm#1{\big|\!\big|#1\big|\!\big|}
\def\Norm#1{\Big|\!\Big|#1\Big|\!\Big|}
\def\normm#1{\bigg|\!\bigg|#1\bigg|\!\bigg|}

\widemargins
\bookheader{NOTES ON ADDITIVE COMBINATORICS}{MARCEL K. GOH}

\maketitle{Notes on additive combinatorics}{by}{Marcel K. Goh}{23 March 2022}

\floattext4.5 \ninebf Note.
\ninepoint
I am compiling this set of expository notes primarily to solidify these fundamental results of
additive combinatorics in
my own mind. Sources that are far more comprehensive exist on the web, so I'm not sure
how useful this will be to anyone else. Currently, the bulk of these notes come from
a series of YouTube lectures of Timothy Gowers (given in early 2022), though I intend
in future to add more relevant topics as I learn about them.

\bigskip

\advsect Pl\"unnecke's theorem

Let $A$ and $B$ be finite subsets of an abelian group. We define the {\it sumset}
$A+B = \{a+b : a\in A,\, b\in B\}$ as well as the {\it difference set}
$A-B = \{a-b : a\in A,\, b\in B\}$. For a nonnegative integer $r$, the {\it $r$-fold iterated sumset}
$rA$ is defined recursively by $0A = \{0\}$ and $rA = A+(r-1)A$. The goal of this first section
is to show that if the cardinality of
a sumset $|A+B|$ is small compared to the size of the set $A$, then the difference
$rB-sB$ of two iterated sumsets must also be small relative to the size of $A$,
in some sense. This result is called Pl\"unnecke's
theorem as it follows from a statement given by H.~Pl\"unnecke in~1969, though this particular
formulation of it was stated and proved by I.~Z.~Ruzsa in 1989.

\edef\thmplunnecke{\the\thmcount}
\parenproclaim Theorem {\advthm} (Pl\"unnecke, {\rm 1969;} Ruzsa, {\rm 1989}). Let $A_0$ and
$B$ be finite subsets of an abelian group and suppose that $|A_0+B|\le K_0|A_0|$ for some constant $K_0$.
Then there exist $A\subseteq A_0$ and $K\le K_0$ such that
for any nonnegative integers $r$ and $s$, not both zero, we have $|rB-sB| \le K^{r+s}|A|$. In particular,
this means that $|rB-sB| \le {K_0}^{r+s}|A_0|$.

Ruzsa used Menger's theorem from graph theory to prove this theorem, but we will use a shorter
argument from a 2012 paper by G.~Petridis. It hinges on the following lemma. In this proof (and the
rest of these notes), we will write $A+x$ instead of $A+\{x\}$ for convenience.

\edef\lempetridis{\the\thmcount}
\parenproclaim Lemma {\advthm} (Petridis, {\rm 2012}). Let $A$ and $B$ be subsets of an abelian
group and let $K$ be such that $|A_0+B| = K_0|A_0|$. Then there exist $A\subseteq A_0$ and
$K\le K_0$ such that $|A+B+C| = K|A+C|$ for every finite subset $C$ of the group.

\proof We pick $A\subseteq A_0$ nonempty such that the ratio $|A+B|/|A|$, which we shall
set as $K$, is minimised. So $|A+B| = K|A|$ and $|A'+B|\ge K|A'|$ for all $A'\subseteq A_0$.
Now we shall show by induction on $C$ that $|A+B+C|\le K|A+C|$ for all $C$. If $C$ is empty we
are done, of course. Now assume the result holds for $C$ and let $x\notin C$ be arbitrary. We have
$$\bigl| A+(C\cup \{x\}) \bigr| = |A+C| + |A+x| - |A'+x|$$
where $A' = \{a\in A : a+x \in A+C\}$. But adding $x$ to everything in a set does not change its
cardinality, so we have
$$|A+C| + |A+x|-|A'+x| = |A+C| + |A| + |A'|.$$
Also, any element of $A'+B+x$ is an element of $A+B+C$, since $a'+x\in A+C$ for all $a'\in A'$.
Applying the induction hypothesis now gives
$$\eqalign{
\bigl| A + B + (C\cup \{x\}) \bigr| &\le |A+B+C| + |A+B+x| - |A'+B+x| \cr
&\le K|A+B| + K|A| = K|A'| \cr
&= K\bigl| A+(C\cup \{x\})\bigr|,
}$$
which completes the proof.\slug

This lemma has an immediate corollary that already has some resemblance to the version of Pl\"unnecke's
theorem that we eventually want to prove.

\proclaim Corollary {\advthm}. If $|A_0 + B|\le K_0|A_0|$, then there is $A\subseteq A_0$ nonempty
and $K\le K_0$ such that $|A+rB| \le K^r |A|$ for all nonnegative integers $r$.

\proof The previous lemma gives $A$ and $K$ such that $|A+B+C|\le K|A+C|$ for all $C$. We now induce
on $r$. The case $r=0$ is obvious. Now for $r>0$ applying the lemma again with $C = (r-1)B$ and
then invoking the induction hypothesis gives
$$|A+rB| = |A+B+(r-1)B| \le K|A+(r-1)B| \le K^r|A|.\noskipslug$$

To get from this corollary to the full theorem, we will use a certain inequality of Rusza (1996), which
is often called the Ruzsa triangle inequality due to its similarity to the triangle inequality in
metric spaces.

\parenproclaim Lemma {\advthm} (Ruzsa triangle inequality). Let $U$, $V$, and $W$ be finite
subsets of an abelian group. Then
$$|U|\cdot |V-W| \le |U-V|\cdot |U-W|.$$

\proof For each $x\in V-W$, pick $v(x)\in V$ and $w(x)\in W$ such that $v(x)-w(x) = x$ (there may
be multiple choices for each of $v(x)$ and $w(x)$, but we arbitrarily pick one).
Define $\phi : U\times (V-W) \to (U-V)\times (U-W)$ by $\phi(u,x) = \bigl(u-v(x), u-w(x)\bigr)$.
If $\bigl(u_1 - v(x_1), u_1 - w(x_2)\bigr) = \bigl(u_2 - v(x_2), u_2 - w(x_2)\bigr)$, then
subtracting the second coordinate of each pair from the first, we find that
$$x_1 = v(x_1) - w(x_1) = v(x_2) - w(x_2) = x_2,$$
and this also implies that $u_1 = u_2$. So $\phi$ gives an injection from $U\times (V-1)$
to $(U-V)\times (U-W)$.\slug

If we define $d(U,V) = |U-V| / \bigl(|U|^{1/2}|V|^{1/2}\bigr)$, then this lemma states
that $d(U,W) \le d(V,U)d(U,W)$. Taking logarithms gives a genuine additive triangle inequality,
but the space of finite subsets of an abelian group does not form a metric space, since
$|U-U|/|U|$ does not equal $1$ in general. In any case, we are now able to prove Ruzsa's
version of Pl\"unnecke's inequality.

\medskip
\noindent{\it Proof of Theorem~{\thmplunnecke}.}\enspace Let $A$ and $K$ be given by
Lemma~{\lempetridis}. By the corollary to that lemma, we have $|A+rB| \le K^r|A|$ and
$|A+sB| \le K^r|A|$. Cardinalities are not changed by taking additive inverses of
everything, so $|{-A}-rB| \le K^r|A|$ and $|{-A}-sB|\le K^r|A|$. Now an application of Ruzsa's triangle
inequality with $U = -A$, $V = rB$ and $W = sB$ yields
$$|A|\cdot|rB-sB| = |{-A}|\cdot|rB-sB| \le |{-A}-rB|\cdot|{-A}-sB| \le K^{r+s}|A|^2,$$
and dividing both sides by $|A|$ proves the theorem.\slug

\advsect Khovanskii's theorem

The goal of this section is to prove the following wonderful theorem of A.~Khovanskii.

\edef\thmkhovanskii{\the\thmcount}
\parenproclaim Theorem {\advthm} (Khovanskii, {\rm 1992}). Let $A$ be a finite
subset of an abelian group and let $f_A(n) = |nA|$ for each $n\in \NN$. Then there exists
$n_A$ and a polynomial $p_A$ such that $f_A(n) = p_A(n)$ for all $n\ge n_A$.

Consider the example when $A = \{0,1,b-1,b\} \subseteq \ZZ$. Then letting $[a,b] = \{x\in \ZZ : a\le x\le b\}$,
we have
$$nA = [0, n] \cup [b-1,\ldots, b+n-1] \cup [2b-2, 2b+n-2] \cup\cdots\cup [n(b-1),nb].$$
This has size $(n+1)^2$ as long as $n< b-1$, but when $n\ge b-1$, $nA$ is simply the set
of integers in the range $[0,bn]$, which has size $bn+1$.

We shall present a combinatorial proof of Khovanskii's theorem,
given by M.~B.~Nathanson and I.~Z.~Ruzsa in 2002.
Let $A = \{a_1,\ldots, a_r\}$ and
for $x = (x_1,\ldots, x_r)\in {\NN_0}^r$, write
$$\sigma(x) = \sum_{i=1}^r x_i\qquad\hbox{and}\qquad
\alpha(x) = \sum_{i=1}^r a_ix_i.$$
Then
$$nA = \{ \alpha(x) : x\in {\NN_0}^r,\, \sigma(x) = n\}.$$
Let $\prec$ be the lexicographic order on ${\NN_0}^r$; that is, $(x_1,\ldots,x_r)\prec (y_1,\ldots,y_r)$
if there is some $i\in [r]$ such that $x_i<y_i$ and $x_j = y_j$ for all $1\le j<i$.
Say that $x$ is {\it useless} if there is $x'\prec x$ such that $\sigma(x') = \sigma(x)$ and
$\alpha(x') = \alpha(x)$; otherwise, say that $x$ is {\it useful}. It is easy to see that
$|nA|$ is exactly the number of useful $x$ with $\sigma(x) = n$. Now let $\le$ denote
the usual partial order on ${\NN_0}^r$, where $(x_1,\ldots,x_r)\le (y_1,\ldots,y_r)$
if $x_i\le y_i$ for all $1\le i\le r$.

\proclaim Lemma {\advthm}. If $x\le y$ and $x$ is useless, then $y$ is useless.

\proof Since $x$ is useless, we can find $x'\prec x$ with $\sigma(x') = \sigma(x)$ and $\alpha(x') = \alpha(x)$.
Let $y' = y+x'-x$. Since $x\le y$, $y-x$ is in ${\NN_0}^r$ and therefore so is $y'$. Then since
$x_i' < x_i$ for the first $i$ at which the two tuples differ, the first coordinate at which $y'$ and
$y$ differ is also $i$ and we have $y_i'<y_i$. Hence $y'\prec y$. Lastly,
$$\sigma(y') = \sigma(y+x'-x) = \sigma(y) + \sigma(x') - \sigma(x) = \sigma(y)$$
and the same is true with $\sigma$ replaced by $\alpha$, so $y$ is useless.\slug

For each useless $x$ we can find a useless $x'$, minimal in the $\le$ order, such that $x'\le x$.
This lemma gives a converse; it says that if $x'$ is a minimal useless element and $x'\le x$, then
$x$ is also useless.

\edef\lemminimalfinite{\the\thmcount}
\proclaim Lemma {\advthm}. For any nonempty $U\subseteq {\NN_0}^r$, the set of minimal elements of
$U$ (with respect to the $\le$ order) is finite.

\proof We perform induction on $r$. If $r=1$, then it is clear that $U$ has exactly one minimal element.
Now let $r>1$ and let $u$ be a minimal element of $u$. Then for every $v\in U$ with $u\not\le v$,
there exists some $i$ and $t< u_i$ such that $v_i = t$. Let
$$V(i,t) = \{v\in U : v_i = t\}$$
for all $1\le i\le r$ and $0\le t<u_i$. Each set $V(i,t)$ is order-isomorphic to a subset of ${\NN_0}^{r-1}$,
so by induction it has finitely many minimal elements. The number of $V(i,t)$ is $r\prod_{i=1}^r u_i$,
and every minimal element of $u$ is either $u$ or a minimal element of some $V(i,t)$, so there are
finitely many minimal elements in $U$.\slug

We will need just one more lemma before we are able to prove Khovanskii's theorem. It shows that
for any fixed $r$-tuple $x'$, the number of $x$ with $x'\le x$ and $\sigma(x) = n$
is given by a polynomial in the variable $n$, with coefficients depending on $x'$ and $r$.

\proclaim Lemma {\advthm}. Let $x'\in {\NN_0}^r$ and for $n\ge \sigma(x')$, let
$$C(x', n) = \{n\in {\NN_0}^r : \sigma(x) = n,\,x'\le x\}.$$
Then
$$\bigl| C(x',n)\bigr| = {n-\sigma(x') + r-1 \choose r-1}.$$

\proof Given $x\in C(x',n)$, map it to $x-x'$. This is a bijection from $C(x',n)$ to $C(0, n-\sigma(x')$.
Now let $m = n-\sigma(x')$. The map $(x_1,\ldots,x_r) \mapsto (x_1,+1,\ldots, x_r+1)$ is a bijection
from $C(0,m)$ to the set of $y\in \NN^r$ such that $y_1+\cdots+y_r = m+r$. Now the map
$$(y_1,\ldots,y_r) \mapsto \{y_1, y_1+y_2, \ldots, y_1+\cdots+y_r\}$$
is a bijection from the previous set to the set of subsets of $[m+r]$ that contain $m+r$. (It is
inverted by the map $\{a_1, \ldots, a_r\} \mapsto (a_1, a_2-a_1, \ldots, a_r-a_{r-1})$.) But this final
set has size ${m+r-1\choose r-1}$, so we are done.\slug

Without further ado, let us now prove Khovanskii's theorem.

\medskip\noindent {\it Proof of Theorem~{\thmkhovanskii}.}\enspace Let $U$ be the set of all
useless $r$-tuples. By Lemma~{\lemminimalfinite}, the set $X\subseteq U$ of all minimal useless
$r$-tuples is finite. Recall that $|nA|$ is the number of useful $x$ such that $\sigma(x) = n$,
so using the notation of the previous lemma, we have
$$|nA| = \bigl| C(0,n)\bigr| - \Bigl| \bigcup_{x'\in X} C(x',n)\Bigr|
=\bigl| C(0,n)\bigr| - \sum_{Y\subseteq X} (-1)^{|Y|} \Bigl| \bigcap_{x'\in Y} C(x',n)\Bigr| $$
But for any $Y\subseteq X$, $\bigcap_{x'\in Y} C(x',n) = C(x_Y, n)$, where $x_Y$ has for its
$i$th coordinate the integer $\max_{y\in Y} y_i$. Hence every term in the inclusion-exclusion
formula depends polynomially on $n$ once
$$n\ge \sum_{1\le i\le r} \max_{x\in X} x_i.\noskipslug$$

\advsect The Balog--Szemer\'edi--Gowers theorem

Let $A$ be a finite nonempty subset of an abelian group. An {\it additive quadruple} in
$A$ is a quadruple $(x,y,z,w)\in A^4$ such that $x+y = z+w$. The {\it additive energy}
$E(A)$ of $A$ is the number of additive quadruples in $A$. If $(x,y,z,w)$ is an additive
quadruple, then $w = x+y-z$, so $E(A)\le |A|^3$. The following lemma roughly states that if
a set has a small sumset, then it has large additive energy.

\proclaim Lemma {\advthm}. Let $A$ be a finite subset of an abelian group and let $C$ be such
that $|A+A|\le C|A|$. Then $E(A)\ge |A|^3/C$.

\proof Define $f : A+A\to \NN$ by letting $f(d)$ equal the number of pairs $(a,b)\in A^2$ such
that $a+b =d$. Then
$$\sum_{d\in A+A} f(d) = |A|^2\qquad\hbox{and}\qquad \sum_{d\in A+A} f(d)^2 = E(A).$$
By the Cauchy--Schwarz inequality,
$$|A|^2 = \sum_{d\ge 1} 1\cdot f(d)
\le \Bigl(\sum_{d\ge 1} 1\Bigr)^{1/2} \Bigl( \sum_{d\ge 1} f(d)^2\Bigr)^{1/2}
\le C^{1/2} |A|^{1/2} E(A)^{1/2}.$$
From here, squaring the whole inequality and then dividing through by $C|A|$ gives the statement
we want to prove.\slug

The converse of this lemma does not hold. For example, we could take the disjoint
union of an arithmetic progression with a geometric progression of the same length.
This set has a large additive energy,
because of the arithmetic progression, but it also has a large sumset, because of the geometric progression.
However, we can pass to a large subset that has a small sumset (namely, the arithmetic progression), and this
is essentially the statement of the Balog--Szemer\'edi--Gowers theorem. We shall state it later on in the
section, after we have given the graph-theoretic definitions and lemmas needed in its proof.

Let $G$ be a bipartite graph with finite vertex sets $X$ and $Y$. For a vertex $v\in X\cup Y$,
let $d(v) = \deg(v)$ denote the {\it degree} of $x$, that is, the number of vertices adjacent to $v$.
For $x_1,x_2\in X$, let $d(x_1,x_2)$ denote the number of paths of length $2$ from $x_1$ to $x_2$. This
is called the {\it codegree} of $x_1$ and $x_2$. The {\it density} of $G$ is the number of edges of $G$
divided by the maximum number of vertices it could have, namely $|X|\cdot|Y|$.

\proclaim Lemma {\advthm}. Let $G$ be a bipartite graph with bipartition $X\cup Y$ and let the density
of $G$ be $\delta$. Then for every $c>0$ there is a subset $X'\subseteq X$ with $|X'|\ge \delta|X|/\sqrt 2$
such that the number of pairs $(x_1,x_2)\in (X')^2$ with $d(x_1,x_2) < c|Y|$ is at most $2c|X|^2/\delta^2$.

\proof Let $y\in Y$ be chosen uniformly at random. Letting $X' = N(y)$, we have
$$\ex\bigl\{ |X'|\bigr\} = \ex_{y\in Y} \bigl\{ d(y)\bigr\} = {1\over |Y|} \sum_{y\in Y} d(y)
= {|E(G)|\over |Y|} = \delta|X|.$$
By convexity of the map $x\mapsto x^2$, we then find that $\ex\bigl\{ |X'|^2\bigr\} \ge \delta^2|X|^2$.
Let
$$B = \bigl\{ (x_1,x_2)\in X^2 : d(x_1,x_2) < c|Y|\bigr\}.$$
The probability that any $(x_1, x_2)\in B$ is also in $(X')^2$ is the same as the probability that
$y$ is connected to both $x_1$ and $x_2$; this is less than $c|Y|$ by the construction of $B$.
So by linearity of expectation, $\ex\bigl\{ |B\cap (X')^2| \bigr\} < c|X|^2$ and in particular, we
have
$$\biggl\{ |X'|^2 - {\delta^2\over 2c} |B\cap(X')^2|\biggr\}
 >\biggl(\delta^2 - {\delta^2\over 2c}\cdot c\biggr)|X|^2 = {\delta^2\over 2}|X|^2.$$
So there is $X'$ such that $|X'| - \bigl(\delta^2/(2c)\bigr) |B\cap (X')^2| \ge \delta^2|X|^2/2$.
This set $X'$ has cardinality at least $\delta |X|/\sqrt 2$ and $|B\cap (X')^2| \le 2c|X'|^2/\delta^2$.\slug

This lemma says that we can find a fairly large subset of $X$ such that almost every pair in this
subset is joined by at least $c|Y|$ paths of length $2$. We would really like to remove the ``almost''
in this statement, but to do so, we will have to pass to paths of length $3$.
First, we need a little lemma about sums of real numbers between $0$ and $1$.

\proclaim Lemma {\advthm}. Let $a_1,\ldots, a_n\in [0,1]$. If $\sum_{i=1}^n a_i \ge an$, then
the number of $i$ such that $a_i\ge b$ is at least
$$ \biggl( {a-b\over 1-b} \biggr) n.$$

\proof Let $I$ be the set of $i$ such that $a_i\ge b$. Then we may write
$$ an \le \sum_{i=1}^n a_i = \sum_{i\in I} a_i + \sum_{i\notin I} a_i
\le |I| + (n-|I|)b.$$
It remains to isolate for $|I|$.\slug

For example, if $b = a/2$, then the number of $i$ with $a_i\ge b$ must be at least $an/2$, and more
generally, if $a= 1-\theta$ and $b=1-\eta$, then the number of such $i$ is at least $(1-\theta/\eta)n$.
If $\eta = 2\theta$, then $a_i\ge 1-2\theta$ at least half the time.
Next we prove the lemma alluded to earlier, concerning paths of length $3$.

\proclaim Lemma {\advthm}. Let $G$ be a bipartite graph with vertex sets $X$ and $Y$ and density $\delta$.
Then there are subsets $X'\subseteq X$ and $Y'\subseteq Y$ such that
$$|X'| \ge {\delta^2\over 8\sqrt 2} |X|\qquad\hbox{and}\qquad
|Y'|\ge {\delta\over 4}|Y|$$
and every $x\in X'$ and $y\in Y'$ are joined by at least $(\delta^6/2^{13}) |X|\cdot |Y|$ paths
of length $3$ in $G$.

\proof By letting $n = |X|$ and setting the $a_i$ equal to the degrees of the vertices in $X$, we have
$a = \delta|Y|$
and we can choose $b=a/2$ in the previous lemma
to deduce that $X$ has a subset $X_1$ such that every vertex in $X_1$ has degree at least
$\delta|Y|/2$ and $|X_1|\ge \delta|X|/2$. Note that the density of the bipartite subgraph with
vertex set $X_1\cup Y$ is $\ge \delta/2$,
so we may invoke the lemma about paths of length $2$ with $\delta$ set to $\delta/2$. It tells us that
for any $c>0$, $X_1$ has a subset $X_2$ of size $\delta|X_1|/2^{3/2}$ such that for all but at most
$8c|X_2|^2/\delta^2$ pairs $(x_1,x_2)\in {X_2}^2$, we have $d(x_1,x_2) \ge c|Y|$.

Now let $\Gamma$ be a graph with vertex set $X_2$ and an edge between $x_1$ and $x_2$ if and only if
$\delta(x_1,x_2)\ge c|Y|$. Note that by this definition, there may be loops in $\Gamma$. In any case,
the average degree in $\Gamma$ is at least $(1-8c/\delta^2)|X_2|$. Invoking the previous lemma again,
at least half the vertices in $X_2$ have degree at least $(1-16c/\delta^2)|X_2|$. Let $X'$ be the set of
all such vertices. Then
$$|X'| \ge {|X_2|\over 2} \ge {\delta|X_1|\over 4\sqrt 2} \ge {\delta^2 |X|\over 8\sqrt 2}.$$
Every vertex in $X_2$ has degree at least $\delta |Y|/2$, so the bipartite subgraph on $X_2\cup Y$
has density at least $\delta/2$ and the average degree of $y\in Y$ is at least $\delta|X_2|/2$. The
previous lemma comes in handy once again, telling us that at least $\delta|Y|/4$ vertices in $Y$ has
degree at least $\delta|X_2|/4$. Let $Y'$ be the set of these vertices.

For any $x\in X'$ and $y\in Y'$, we know that $x$ has at least $(1-16c/\delta^2)|X_2|$ neighbours in $\Gamma$
and $y$ has at least $\delta|X_2|/4$ neighbours in $X_2$. There must be at least
$(\delta/4-16c/\delta^2)|X_2|$ neighbours in commmon. Letting $c=\delta^3/128$, the number of neighbours
in common becomes $\delta|X_2|/8$, and the number of paths of length $3$ from $x$ to $y$ is at least
$$ {\delta |X_2|\over 8} |X_2| \cdot c|Y| = {\delta\over 8}\cdot {\delta^3\over 128}
\cdot {\delta^2\over 4\sqrt 2} |X|\cdot |Y| \ge {\delta^6\over 2^{13}}|X|\cdot |Y|,$$
as foretold.\slug

Let $A$ be a finite subset of an abelian group. We say that an element $d\in A-A$ is {\it $\theta$-popular}
if $\bigl|\bigl\{ (a,b)\in A^2 : b-a=d\bigr\}\bigr|\ge \theta |A|$.

\proclaim Lemma {\advthm}. Let $A$ be a finite subset of an abelian group and let $E(A)$ denote
the additive energy of $A$. If $E(A)\ge c|A|^3$, then there are at least $c|A|/2$ elements of $A-A$
that are $(c/2)$-popular.

\proof Define $f : A-A\to \NN$ by $f(d) = \bigl|\bigl\{ (a,b)\in A^2 : b-a=d\bigr\}\bigr|$.
Note that
$$\sum_{d\in A-A} f(d) = |A|^2\qquad\hbox{and}\qquad \sum_{d\in A-A} f(d)^2 = E(A).$$
Let $P$ be the set of $(c/2)$-popular differences and let $U = (A-A)\setminus P$. Then
$$\eqalign{
c|A|^3 &\le E(A)\cr
&= \sum_{d\in A-A} f(d)^2 \cr
&= \sum_{d\in P} f(d)^2 + \sum_{d\in U} f(d)^2 \cr
&\le |P|\cdot |A|^2 + {c|A|\over 2} \sum_{d\in U} f(d) \cr
&= |P|\cdot |A|^2 + {c|A|^3\over 2}, \cr
}$$
which implies that $|A|\ge c|A|/2$.\slug

We are, at long last, ready to state and prove the Balog--Szemer\'edi--Gowers theorem. It is
named for A.~Balog and E.~Szemer\'edi, who first proved it in 1994, as well as for W.~T.~Gowers,
who gave an alternative proof in 1998 that yielded power-type bounds, a vast improvement to the tower-type
bounds in Balog and Szemer\'edi's proof.

\parenproclaim Theorem {\advthm} (Balog--Szemer\'edi, {\rm 1994;} Gowers, {\rm 1998}).
Let $A$ be a finite subset of an abelian group with $E(A)\ge c|A|^3$. Then $A$ has a subset
$A'$ of size at least $c'|A|$ such that $|A'-A'|\le C|A'|$, where $c'$ and $C$ have a power-type
dependence on $c$. In particular, we may take
$$ c' = {c^4\over 2^8}\qquad\hbox{and}\qquad C = {2^{68}\over c^{36}}.$$

\proof Define a bipartite graph $G$ with both vertex sets being copies of $A$ (though
we'll call one of these copies $B$ to avoid any confusion). Join $a\in A$ to
$b\in B$ if and only if $b-a$ is $(c/2)$-popular. Each $(c/2)$-popular difference leads to at least
$c|A|/2$ edges, so by the lemma, there are at least $c^2|A|^2/4$ edges. Let $\delta = c^2/4$.
By the lemma concerning paths of length $3$, there exist $A'\subseteq A$ and $B'\subseteq B$
with
$$ |A'| \ge {\delta^2\over 8\sqrt 2} |A|\qquad\hbox{and}\qquad
|B'|\ge {\delta \over 4} |A|$$
such that every $a\in A$ and $b\in B$ are joined by at least $\delta^6|A|^2/2^{13}$ paths of length $3$
in $G$.

Given a path $(a,u,v,b)$ of length $3$, we have $b-a = u-a - (u-v) + b-v$, where $u-a$, $u-v$, and $b-v$
are all $(c/2)$-popular. Hence the number of ways of writing $b-a$ as $r_1-s_1 - (r_2-s_2) + r_3 - s_3$
is at least
$${\delta^6\over 2^{13}} |A|^2 \biggl( {c\over 2} |A|\biggr)^3 = {c^{15}\over 2^{28}} |A|^5.$$
But the number of possible $(r_1,s_1,r_2,s_2,r_3,s_3)$ is $|A|^6$. So
$$ |B'-A'|\cdot {c^{15}\over 2^{28}} |A|^5 \le |A|^6 $$
and $|B'-A'|\le 2^{28}|A|\over c^{15}$. Now the Ruzsa triangle inequality with $U = B'$ and
$V = W = A'$ tells us that $|B'|\cdot |A'-A'| \le |B'-A'|^2$, so
$$|A'-A'|\le {|B'-A'|^2\over |B'|} \le {2^{56}|A|^2\over c^{30}}\cdot {2^4\over c^2|A|}={2^{60}\over c^{32}}|A|.$$
But recall that
$$|A'| \ge {\delta^2\over 8\sqrt 2} |A| \ge {c^4\over 2^8}|A|,$$
so $|A'-A'| \le 2^{68}|A'|/c^{36}$.\slug

\advsect Basics of discrete Fourier analysis

Let $Z$ be a finite abelian group. A {\it character} on $Z$ is a homomorphism from
$Z$ to the multiplicative group $\C\setminus\{0\}$. If $\chi$ is such a homomorphism, then
$\bigl|\chi(x)\bigr| = 1$ for every $x\in Z$, so we can actually regard $\chi$ as being
a function from $Z$ to the unit circle $\TT = \{z\in \C : |z| = 1\}$. The pointwise product
of two characters gives another character, and the function $\chi_0$ that sends every $x\in Z$
to $1$ has the property that if $\chi$ is any character, then $\chi\chi_0 = \chi = \chi_0\chi$.
Lastly, we note that multiplication is commutative and
for any character $\chi$, the product of $\chi$ with $\bar\chi$
gives $\chi_0$, so the set of characters on $Z$ is an abelian group. This is called the
{\it dual group} or sometimes {\it Pontryagin dual} of $Z$ and is denoted $\hat Z$.

We define an inner product on the space $\CC^Z$ of functions from $Z$ to $\CC$ by setting
$$\langle f,g\rangle = \ex_x f(x) \bar{g(x)},$$
where $\ex_{x\in Z} F(x) = |Z|^{-1} \sum_{x\in Z} F(x)$.
We can also make $\hat Z$ into an inner product space by letting
$$\langle \hat f, \hat g\rangle = \sum_{\chi\in \hat Z} \hat f(\chi)\bar{\hat g(\chi)};$$
note that this time we do not normalise by dividing by $|Z|$.
Two functions $f$ and $g$ in an inner product space are said to be {\it orthogonal} if $\langle f,g\rangle = 0$.
The first lemma we'll prove concerns certain orthogonality relations.

\parenproclaim Lemma {\advthm} (Orthogonality relations). Let $Z$ be a finite abelian group.
\medskip
\item{a)} If $\chi_1$ and $\chi_2$ are characters on $Z$, then
$$\langle \chi_1,\chi_2 \rangle = \ex_{x\in Z} \chi_1(x)\bar{\chi_2(x)}
  = \cases{ 1, & if $\chi_1 = \chi_2$;\cr 0, & if $\chi_1 \ne\chi_2$.}.$$
\smallskip
\item{b)} For $x,y\in Z$, we have
$$\sum_{\chi\in \hat Z} \chi(x)\bar{\chi(y)}
  = \cases{ |\hat Z|, & if $x=y$;\cr 0, & if $x\ne y$,}$$
\medskip

\proof Let $\chi_0$ denote the trivial character. We have
$$\ex_{x\in Z} \chi_0(x) = \ex_{x\in Z} 1 = 1.$$
On the other hand, let $\chi$ be a nontrivial character and let $u\in Z$ be such that $\chi(u)\ne 1$. Then from
$$\ex_{x\in Z} \chi(x) = \ex_{x\in Z} \chi(u+x) = \ex_{x\in Z}\chi(u)\chi(x) = \chi(u)\ex_{x\in Z}\chi(x),$$
it follows that $\ex_{x\in Z}\chi(x) = 0$. Now consider $\chi_1\bar{\chi_2}$. If $\chi_1 = \chi_2$, then
this is the trivial character and from our first observation, $\langle \chi_1,\chi_2\rangle = 0$. Otherwise,
we are in the second case and the inner product is zero. This proves part (a).

Part (b) is proven similarly. Note that $\chi(x)\bar{\chi(y)} = \chi(x-y)$. If $x=y$, then we have
$$\sum_{\chi\in\hat Z} \chi(x-y) = \sum_{\chi\in\hat Z} \chi(0) = |\hat Z|.$$
If $x\ne y$, then there is some $\psi\in \hat Z$ such that $\psi(x-y)\ne 1$, and then writing
$$\sum_{\chi\in \hat Z} \chi(x-y) = \sum_{\chi\in\hat Z} \psi(x-y)\chi(x-y)
= \psi(x-y)\sum_{\chi\in\hat Z} \chi(x-y),$$
we see that $\sum_{\chi\in\hat Z} \chi(x-y)$ must be zero.\slug

Since the space of functions from $Z\to \CC$ has dimension $|Z|$, there are at most $|Z|$ characters.
Every finite abelian group can be written as a direct product of cyclic groups, and we shall use this fact
to show that there are exactly $|Z|$ characters; i.e., $|\hat Z| = |Z|$. For brevity, let $\ZZ_n$ stand for $\ZZ/n\ZZ$ for all $n\in\NN$,
and for $\alpha\in \RR$, let $e(\alpha) = e^{2\pi i\alpha}$.

\proclaim Lemma \advthm.
The set of characters on a finite abelian group $Z$ spans the space of functions from $Z$ to $\CC$.

\proof Write $Z \cong \ZZ_{n_1} \times \ZZ_{n_2}\times \cdots\times \ZZ_{n_r}$. For every
$u = (u_1,\ldots, u_r)\in Z$, the function $\chi_u : Z\to \CC$ that takes
$$(x_1,\ldots, x_r) \mapsto \prod_{i=1}^r e\biggl( {u_ix_i\over n_i}\biggr)$$
is a character (this is easy to show, since $e(x_i + x_i') = e(x_i)e(x_i')$). If $u\ne u'$,
then $\chi_u\ne\chi_{u'}$ and $\langle \chi_u,\chi_{u'}\rangle = 0$ by the previous lemma. Hence
the set $\{\chi_u : u\in Z\}$ comprises $|Z|$ linearly independent elements in a space of dimension $|Z|$,
which must be spanning.\slug

Note that the map $u\mapsto \chi_u$ gives an isomorphism from $Z$ to $\hat Z$.
The two main examples we shall consider are when $Z = \ZZ_n$ and when $Z = \FF_p^n$. In the first
case, $\chi_u$ takes $x\mapsto e(ux/N)$ and in the second case $\chi_u$ takes $x\mapsto e(u\cdot x/p)$.
(Recall that if $u = (u_1,\ldots, u_n)$ and $x = (x_1,\ldots, x_n)$ then the dot product $u\cdot x$
is the sum $\sum_{i=1}^n u_ix_i$.)

For a finite set $X$, we shall denote by $L_p(X)$ the normed vector space of all functions $f:X\to \CC$
under the norm
$$\norm{f}_p = \bigl( \ex_{x\in X} \bigl| f(x)\bigr|^p \bigr)^{1/p}.$$
The notation $l_p(X)$ denotes the same set, but with the norm
$$\norm{f}_p = \Bigl( \sum_{x\in X} \bigl| f(x)\bigr|^p \Bigr)^{1/p}$$
instead. Note that $\norm{f}_2^2 = \langle f,f\rangle$ and $\norm{f}_\infty = \max_{x\in X} \bigl|f(x)\bigr|$.

\medskip\boldlabel The Fourier transform.
If $f:Z\to \CC$, the {\it Fourier transform} of $f$ is the function $\hat f : \hat Z\to\CC$ given by
$$\hat f(\chi) = \langle f,\chi\rangle = \ex_{x\in Z} f(x) \bar{\chi(x)}.$$
We shall now state and prove an identity which is extremely useful in Fourier analysis.

\parenproclaim Lemma {\advthm} (Parseval's identity). Let $Z$ be a finite abelian group
and let $f$ and $g$ be functions from $Z$ to $\CC$.
If $\hat f$ and $\hat g$ are the Fourier transforms of $f$ and $g$ respectively, then
$\langle f,g\rangle = \langle \hat f, \hat g\rangle$.

\proof First, we expand all the definitions and rearrange sums to obtain
$$\eqalign{
\langle \hat f, \hat g\rangle &= \sum_{\chi\in \hat Z} \hat f(\chi)\bar{\hat g(\chi)} \cr
&=  \sum_{\chi\in \hat Z} \ex_{x\in Z} f(x)\bar{\chi(x)} \ex_{y\in Z} g(y)\bar{\chi(y)} \cr
&=   \ex_{x\in Z} f(x) \ex_{y\in Z} g(y) \sum_{\chi\in \hat Z}\bar{\chi(x)}\bar{\chi(y)}.\cr
}$$
By the second orthogonality relation and the fact that $|\hat Z| = |Z|$, the inner sum equals $|Z|$ when
$y = x$ and $0$ otherwise. Thus for any given $x$, we have
$$\ex_{y\in Z} g(y) \sum_{\chi\in \hat Z}\chi(x)\bar{\chi(y)} = g(x),$$
which is exactly what we need for the whole thing to equal $\langle f,g\rangle$.\slug

We also have the following inversion formula that recovers the original function $f$ from $\hat f$.

\parenproclaim Lemma {\advthm} (Fourier inversion formula). Let $Z$ be a finite abelian group and
let $f:Z\to \CC$. We have
$$f(x) = \sum_{\chi\in \hat Z} \hat f(\chi)\chi(x).$$

\proof We expand
$$\sum_{\chi\in \hat Z} \hat f(\chi)\chi(x) = \sum_{\chi\in \hat Z} \ex_{y\in Z} f(y) \bar{\chi(y)} \chi(x)
= \ex_{y\in Z} f(y) \sum_{\chi\in \hat Z} \chi(x)\bar{\chi(y)},$$
and we noted in the proof of Parseval's theorem that the right-hand side is exactly $f(x)$.\slug

The inversion formula tells us that two functions from $Z\to \CC$ are equal if and only if their Fourier
transforms are equal. For the next lemma, which concerns dilates of a function $f: Z\to \CC$, we will
establish some notation. If $x\in Z$, we define $nx$ recursively for all integers $n\ge 0$ by $0x = 0$
and $nx = x + (n-1)x$. We then set $nz = - (-n)x$ for all negative integers. Since the group operation
on characters is multiplication, we will not write $n\chi$ but instead $\chi^n$ for $\chi$ multiplied
with itself $n$ times.

\parenproclaim Lemma {\advthm} (Dilation rule). Let $Z$ be a finite abelian group and let $a\in \ZZ$
be an integer that is coprime to $|Z|$. Denote the multiplicative inverse of $a$ modulo $|Z|$ by $a^{-1}$.
Letting $f_a : Z\to \CC$ be the function given by $f_a(x) = f(a^{-1} x)$, we have
$\hat{f_a}(\chi) = \hat f(\chi^a)$.

\proof We have
$$\hat{f_a}(\chi) = \ex_{x\in Z} f_a(x) \bar{\chi(x)} = \ex_{x\in Z} f(a^{-1} x)\bar{\chi(x)}.$$
Since $x\mapsto ax$ is a bijection from $Z$ to itself, we can replace $x$ formally by $ax$ in the above
to get
$$\hat{f_a}(\chi) = \ex_{x\in Z} f(x)\bar{\chi(ax)} = \ex_{x\in Z} f(x) \bar{\chi(ax)}^a
= \hat f(\chi^a).\noskipslug$$

\medskip\boldlabel Convolutions.
For functions $f$ and $g$ from a finite abelian group $Z$ to $\CC$, the {\it convolution} $f*g$ is
defined by
$$ (f*g)(x) = \ex_{y+z = x} f(y)g(z).$$
If instead $\hat f$ and $\hat g$ are functions from $\hat Z$ to $\CC$, then
$$ (\hat f * \hat g)(\chi) = \sum_{\chi_1\chi_2 = \chi} \hat f(\chi_1)\hat g(\chi_2).$$

\parenproclaim Lemma {\advthm} (Convolution law). Let $Z$ is a finite abelian group and $f,g : \Z\to \CC$.
For all $\chi\in \hat Z$, we have $\hat{f*g} (\chi) = \hat f(\chi)\hat g(\chi)$.

\proof We start by expanding
$$\hat{f*g}(\chi) = \ex_{x\in Z} (f*g)(x) \bar{\chi(x)} = \ex_{x\in Z} \ex_{y+z=x} f(y)g(z)\bar{\chi(y)\chi(z)}.$$
But note that $x$ does not appear in the summand anymore, meaning that we can simply rewrite this as
an expectation
over {\it all} $y$ and $z$ (their sum will equal $x$ for some $x\in Z$). Thus we can conclude that
$$\hat{f*g}(\chi) = \ex_{y\in Z}\ex_{z\in Z} f(y)g(z)\bar{\chi(y)\chi(z)} = \hat f(\chi)\hat g(\chi),$$
which is what we wanted to show.\slug

If $A$ is a subset of a finite abelian group $Z$, we associate to $A$ the {\it characteristic function}
$\one_A : Z\to \CC$ given by
$$\one_A(x) = \cases{ 1, & if $x\in A$;\cr 0, & otherwise.}$$
When it will not cause confusion, we will abuse notation and write $A(x)$ instead of $\one_A(x)$.
With the definitions and results stated above, we are now in a position to translate properties of $A$ to
Fourier-analytic statements about the function $\one_A$. First off, note that
$$\bignorm{\hat{\one_A}}_2^2 = \norm{\one_A}_2^2 = \ex_{x\in Z} A(x)^2 = \ex_{x\in Z} A(x)
= {|A|\over |Z|}.$$
If $|A|/|Z| = \delta$, then we will say that $A$ has {\it density} $\delta$. Another way of expressing $\delta$
in terms of $\one_A$ is
$$\hat{\one_A}(\chi_0) = \ex_{x\in Z} A(x)\chi_0(x) = \ex_{x\in Z} A(x) = \delta,$$
where $\chi_0$ is the trivial character.

The fact that $\norm{\one_A}_2^2 = \delta$ means that if we sample $x$ and $y$ from $Z$ and
condition that $x=y$, the probability that $x$ (or $y$) is in $A$ is $\delta$. The reason for stating this
obvious fact in such a stilted fashion is that it generalises by way of the convolution operation we defined
earlier; that is, what if we are interested in the probability that a $4$-tuple $(x,y,z,w)$ satisfying $x+y=z+w$
is a member of $A^4$? Well, the convolution law gives
$$\eqalign{
\ex_{x+y=z+w} A(x)A(y)A(z)A(w) &= \ex_{u\in Z} \ex_{x+y=u}\ex_{z+w=u} A(x)A(y)A(z)A(w) \cr
&= \ex_{u\in Z} (\one_A*\one_A)(u)^2 \cr
&= \norm{\one_A*\one_A}_2^2 \cr
&= \bignorm{\hat{\one_A*\one_A}}_2^2 \cr
&= \bignorm{\bigl(\hat{\one_A}\bigr)^2}_2^2 \cr
&= \sum_{\chi\in \hat Z} \bigl| \hat{\one_A}(\chi)\bigr|^4 \cr
&= \norm{\one_A}_4^4, \cr
}$$
and more generally, the probability that a tuple $(a_1,\ldots,a_k, b_1,\ldots,b_k)\in Z^{2k}$
with $a_1+\cdots+a_k=b_1+\cdots+b_k$ is a member of $A^{2k}$ is $\norm{\one_A}_{2k}^{2k}$. Recall that
the number of $(x,y,z,w)\in A^4$ with $x+y = z+w$
is the additive energy $E(A)$ of $A$; we just showed above that
$E(A) = |Z|^3 \norm{\one_A}_4^4$.

\section References

\bye