### 2022-04-13 misc/expo/tensorrank.tex

\input fontmac
\input mathmac

\def\FF{{\bf F}}
\def\TT{{\bf T}}
\def\bar{\overline}
\def\hat{\widehat}
\def\norm#1{|\!|#1|\!|}
\def\bignorm#1{\big|\!\big|#1\big|\!\big|}
\def\Norm#1{\Big|\!\Big|#1\Big|\!\Big|}
\def\normm#1{\bigg|\!\bigg|#1\bigg|\!\bigg|}

\widemargins
\bookheader{THE ANALYTIC RANK OF A TENSOR}{MARCEL K. GOH}

\maketitle{The analytic rank of a tensor}{by}{Marcel K. Goh}{10 April 2022}

\bye